\chapter*{Acknowledgments}
My educational journey until this point has been fruitful and very
memorable, thanks to all the wonderful people who have been a part
of my journey. First and foremost, I am grateful to my advisor,
Prof. Onur Mutlu, who was willing to take me on as his student
despite my lack of background in computer architecture then and
give me the mentorship, time and encouragement to build
background and grow as a researcher. Onur's emphasis on clarity in
thinking, speaking and writing has been a major influence in
shaping me. I am also very thankful to Onur for providing me with
the resources and freedom to carry out research and for always finding
the right opportunities by way of collaborations and internships
to further my research.

I would like to thank my committee members, Prof. Greg Ganger,
Prof. James Hoe and Dr. Ravi Iyer for their time, effort and
inputs in bringing this dissertation to completion. Special thanks
to James for his encouragement and feedback even since my early
years at Carnegie Mellon University. Thanks to Greg for his incisive insights on various
aspects of my work, from his unique perspective as the storage QoS
expert on my committee. Thanks to Ravi for his many inputs and
insights on QoS and for giving me the opportunity to intern with
his spirited and warm group in Intel Labs, Hillsboro. 

The SAFARI group has been a great source of critical feedback,
ideas and fun. I am incredibly thankful for everything I've
learned from this group of smart, enthusiastic and hard-working
graduate students over the years. Vivek Seshadri has been a great
friend and lab mate. He has been an amazing sounding board for new
ideas. Several of the ideas in this thesis have evolved a lot
through numerous discussions with Vivek. Thanks a lot to Yoongu
Kim for all his feedback on writing and presentation. Yoongu's
high standards for research, presentation and writing are an
inspiration. I am thankful to Chris Fallin for his critical
feedback on research and for all I learned from him during the
many times we worked together on TAing, course work, quals. Thanks
a lot to Samira Khan for the many discussions on research and
life, in general. Donghyuk Lee's DRAM expertise and drive to keep
learning are admirable. Kevin Chang's methodical approach to
research and problem solving have been very useful in many of the
projects we have worked together on. Thanks to Rachata
Ausavarungnirun for his helpful nature. Many thanks to visiting
researchers Hui Wang, Hiroyuki Usui and interns Harsha Rastogi,
Arnab Ghosh for working with me on different research projects.
Thanks to Gennady Pekhimenko for being a great cube mate. His work
ethic and discipline are admirable. Thanks to Justin Meza,
Hongyi Xin, Nandita Vijayakumar, Yang Li, Yixin Luo, Kevin Hsieh
and Amirali Bouramand for the many discussions and dinners. 

Besides members of the SAFARI group, several graduate students
have been a great source of advice and encouragement at several
points. Thanks to Siddharth Garg for getting me inducted into the
workings of graduate school and the sound advice, when I was still
learning the ropes as a first year graduate student. Thanks to
Michael Papamichael for all the inputs, advice and discussions
over the years at Carnegie Mellon University. Michael's genuine
passion for research and his ability to explain concepts so
clearly are inspiring. Thanks to Karthik Lakshmanan for the many
discussions and inputs during my early years. Thanks to Anagha
Kulkarni for her encouragement and the many discussions we have
had about graduate school and life in general when at MSR,
Redmond. Thanks to Michelle Goodstein for her company and her
perspectives on life. Many thanks to Elaine Lawrence, Samantha
Goldstein, Karen Lindenfelser, Nathan Snizaski, Marilyn Patete,
Debbie Scappatura and Olivia Vadnais for helping me navigate
through administrative aspects. Thanks also to the CMU shuttle and
escort drivers who have safely ferried me home several late
nights.

I enjoyed my internships at Intel Labs, Hillsboro and Microsoft
Research, Redmond. Thanks to Li Zhao for being a very hands-on and
involved mentor during my internship at Intel Labs. Li was always
available to brainstorm and discuss. Thanks to Trishul Chilimbi,
Sriram Sankar and Kushagra Vaid for being great mentors during my
internship at MSR. Thanks to Thomas Moscibroda for the weekly
brainstorming sessions when I was at MSR. Many thanks to Gabriel
Loh for his mentorship. I would like to thank National Science
Foundation (NSF), Semiconductor Research Corporation (SRC),
Gigascale Systems Research Center (GSRC) and Intel for generously
supporting my research over the years and Carnegie Mellon
University for supporting me with the John and Claire Bertucci
fellowship

I am very thankful to all my teachers from school and professors
from undergrad for instilling in me a basic sense of curiosity and
an urge to learn. Special thanks to Ms. Jennifer, Ms. Nargis, Ms.
Bhuvaneswari, Prof. Mala John and Prof. Kannan. Thanks also to
several seniors from my undergraduate institution who served as
role models and a source of inspiration. I am grateful to Dr.
Sasikanth Avancha from Intel for mentoring me on my undergraduate
project and my manager at SanDisk, Radhakrishnan Nair, for his
support when applying to graduate school.

Graduate school is a long and intense journey, with many highs and
lows. I am very thankful for the support systems I had through
grad school, in terms of friends, room mates and family. Thanks to
Anusha Venkatramani, Ashwati Krishnan, Aishwarya Sukumar, Lavanya
Iyer, Swati Sarraf and Manali Bhutiyani for being great and very
understanding room mates and friends. Thanks to Abhay Ramachandra,
Aditi Pandya, Arun Kannawadi, Arvind Muralimohan, Athula
Balachandran, Bhavana Dalvi, Divya Hariharan, Divya Sharma, Janani
Mukundan, Madhumitha Ramesh, Mahesh Joshi, Natasha Kholgade,
Niranjini Rajagopal, Ramkumar Krishnan, Siddharth Komini Babu,
Siddharth Gopal, Suchita Ramachandran, Swaminathan Ramesh, Varoon
Shankar for their companionship through different points of grad
school. Thanks to Aishwarya R, Anusha Radhakrishnan, Ramya
Guptha, Gayatri Singaravelu, Mukund Kalyanaraman, Sudharsan
Seshadri and Arunachalam Annamalai for their friendship and
support from afar. 

My family has been a big source of support, encouragement and
comfort all through my journey. Thanks to my mother, Bhuvaneswari
and my father, Subramanian for instilling in me the value of
education. My mother, a math teacher, inculcated in me an interest
in math from a young age and has been a big source of
encouragement. My father's work ethic has been an inspiration. My
grand parents have been extremely supportive of all my endeavors,
right from a very young age and I am grateful to them for all the
warmth, support and care over the years. My grand father
Rajamony's enthusiasm until his very last breath was infectious.
My grand mother Jambakam's appreciation of and strong belief in
the need for a solid education, despite her not having had access
to it is admirable. I am also very grateful to my paternal grand
mother Visalakshi for her encouragement - she, for some reason,
thought I was going to be an engineer since I was ten years old. I
wish she had seen me finish my PhD.

My uncle, Kumar, has been a great source of encouragement, knowledge and
fun. He introduced me to science through fun experiments at home.
He introduced me to the wonderful world of books and libraries
from a very young age. He has had an immense role in shaping me
during my formative years. My sister, Lakshmi, has been a pillar of support.
Through school and undergrad, she was always willing and eager to
get me any books and resources I needed. She has been extremely
supportive and involved in all my major academic/career decisions.
My parents-in-law and sister-in-law, Devi have been a big source of
encouragement over the past couple of years.

Finally, Kaushik Vaidyanathan has been a rock solid pair of
dependable shoulders, over the past several years as my best
friend and more recently, my husband. Right from pushing me to
apply to grad school to listening to my long rants during times of
self doubt to putting up with my erratic schedules to what not,
his support has been immense and has made this whole journey
possible and that much more fun and enjoyable. I will not even
attempt to say thank you for that will trivialize all that he has
done for me.
