

\documentclass{article}
\usepackage[accepted]{icml2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\def\UrlBreaks{\do\/\do-}

\usepackage{microtype}
\usepackage{makecell}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{chngcntr}
\usepackage{graphicx}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{scrextend}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage[explicit]{titlesec}
\usepackage{sidecap}

\definecolor{amaranth}{rgb}{0.7, 0.17, 0.31}
\definecolor{airforceblue}{rgb}{0.30, 0.44, 0.56}
\definecolor{darkgreen}{rgb}{0.13, 0.45, 0.13}
\definecolor{bronze}{rgb}{0.8, 0.5, 0.2}

\graphicspath{ {./figs/} }


\newcommand\todo[1]{\textcolor{red}{TODO(#1)}}
\newcommand\pegasusbase{\xspace}
\newcommand\transformerbase{\xspace}
\newcommand\pegasuslarge{\xspace}
\newcommand\pegasuslargegnews{\makecell{\pegasuslarge\C4)} & \makecell{\pegasuslarge\coverage=\frac{1}{S}\sum_{f\in F(A,S)}|f|density=\frac{1}{S}\sum_{f\in F(A,S)}|f|^2ID \#187)} & 
GP practices are being paid to help local NHS groups limit the number of patient referrals and cut costs, the doctors' magazine Pulse found. Appointments affected include scans and consultations with specialists - including those for cancer patients. The British Medical Association said such incentives were "misguided". At least nine clinical commissioning groups (CCGs) were offering GP practices payments for hitting targets, according to Pulse's investigation. In one case, Birmingham South Central CCG was offering practices more than   £11,000 to reduce new outpatient attendances, follow-ups, A\&E attendances and emergency admissions by 1\%, compared with 2014/15. It said the schemes, which exclude cancer referrals, were designed to "incentivise best quality practice" and "drive improvements in the quality of primary medical care". "Our priority is to ensure that patients have access to services that they need, when they need them," said a spokesperson for Birmingham South Central CCG. Another CCG told Pulse it had considered the "full impact" of the incentive scheme and was "confident that there is no conflict of interest". Pulse said that one scheme had already been looked at by the General Medical Council, the body which regulates medical standards in the UK, after local GP leaders expressed their concern. The magazine pointed out that initial hospital referrals for cancer patients should happen within two weeks of a GP first suspecting the condition. Dr Chand Nagpaul, chairman of the GPs committee of the doctors' trade union the British Medical Association, told BBC Radio 4's Today programme that such schemes were a "financial contaminant" to patient-doctor trust. He said: "It's short-sighted and misguided of CCGs to introduce such mechanisms, because they do lead to the potential for patients questioning the motives of GP referrals. "We believe it is far more appropriate for CCGs to introduce clinical pathways that ... ... ... 
\\ \hline
Gold & 
Some doctors in England are being offered thousands of pounds to cut the number of patients being sent to hospital, an investigation has found.
\\ \hline
Model & 
Thousands of pounds are being paid to GPs in England to avoid sending patients to hospital, an investigation has found.
\\ \hline
ROUGE1-F1 & 
68.18
\\ \hline
\makecell{Document\ID \#276)} & 
Thieves took the animal from Silver Star Pets in Pengam, along with dog leads, dog coats, fish tanks, pet food and dog grooming equipment worth £35,000. The incident is one of four break-ins that occurred at the Britannia Enterprise Centre some time between Tuesday evening and Wednesday morning. Police are appealing for information. The other units broken into were Auto Stitch, George Monumental Masons and Pride and Groom where damage was caused to the doors but nothing was stolen. A taxi sign, however, was also stolen from one of the vehicles parked at the centre.
\\ \hline
Gold & 
An African Grey parrot valued at £900 has been stolen from a pet shop in Caerphilly.
\\ \hline
Model & 
A Yorkshire terrier worth £100,000 has been stolen from a pet shop in Stoke-on-Trent.
\\ \hline
ROUGE1-F1 & 
48.48
\\ \hline
\makecell{Document\ID \#124)} & 
Johnson made his film debut in the 1950s and featured in numerous films, alongside stars such as Frank Sinatra, Laurence Olivier and Charlton Heston. A founder member of the Royal Shakespeare Company, Johnson played several lead roles including Romeo and Mark Anthony in Julius Caesar. He also appeared in several TV dramas such as Lewis and Silent Witness. Johnson died after a short illness at the Royal Marsden Hospital in Chelsea, London. He is survived by his wife Lynne, who he married in 2004, and his four children. Johnson was born in Upminster, Essex, and he left his training at the Royal Academy of Dramatic Art (RADA) to join Sir John Gielgud's company. He joined the Royal Navy during World War Two and then made his film debut in 1959, when he appeared in the MGM film Never So Few, starring Frank Sinatra and Gina Lollobrigida. He also appeared in The Haunting (1963) and Khartoum (1966), opposite Laurence Olivier and Charlton Heston. His family said he was offered and turned down the role of James Bond after playing British spy Bulldog Drummond in Deadlier Than the Male (1967) and its sequel Some Girls Do (1969). His most recent film credits include The Boy in the Striped Pyjamas (2008) and Lara Croft: Tomb Raider (2001).
\\ \hline
Gold & 
British actor Richard Johnson, whose career spanned film, theatre and TV, has died aged 87, his family has said.
\\ \hline
Model & 
Actor Roger Johnson, best known for his roles in Deadlier Than the Male and Some Girls Do, has died at the age of 93.
\\ \hline
ROUGE1-F1 & 
27.91
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on XSum sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{XSum} \\ \hline
\makecell{Document\ID \#7)} & 
Operation Anagram was formed by Strathclyde Police in 2006 after Tobin raped and murdered 23-year-old Polish student Angelika Kluk in Glasgow. It also helped convict him of the murders of 18-year-old Dinah McNicol and 15-year-old Vicky Hamilton. Police said that while the operation was being scaled back it would never end. Det Sup David Swindle, who formed the operation, said anyone with information regarding Peter Tobin would still be able to contact the Anagram incident room via e-mail which would be monitored daily. "Since Operation Anagram was formed in 2006, I could never have believed the momentum and magnitude it would have developed into," he said. "The public response and assistance from the media has been unbelievable. "The long-running investigation has brought some solace to the families of Vicky Hamilton and Dinah McNicol in knowing what happened to their loved ones." Det Sup Swindle said that although Anagram had not uncovered other murder victims of Tobin to date, he was "proud" of its "success in bringing closure" to some bereaved families. He added: "I leave Strathclyde Police after 34 years in the knowledge that we have tried our best to find out the truth of what Tobin has done and also with confidence that the Anagram processes built up over the last four and a half years ensures a life long awareness of this cowardly, vicious serial killer Tobin. "Any new information coming to light will continue to be researched which may bring similar closure to other families. "The search for the truth about what Peter Tobin has done will continue." Tobin is serving three life sentences for the murders of Vicky Hamilton, Dinah McNicol and Angelika Kluk. Ms Kluk was raped and murdered at a church in Glasgow in 2006. The bodies of Ms McNicol, from Essex, and Ms ... ... ... 
\\ \hline
Gold & 
A dedicated police operation set up to examine the activities of serial killer Peter Tobin is to be wound down.
\\ \hline
Model & 
A police investigation into serial killer Peter Tobin is being scaled back.
\\ \hline
ROUGE2-F1 & 
26.67
\\ \hline
\makecell{Document\ID \#234)} & 
She will play Denker, a lady's maid to Dame Maggie Smith's character, the Dowager Countess of Grantham. Johnston, who has also appeared in Waking the Dead and Coronation Street, joins new stars Richard E Grant and Anna Chancellor, both of whom will play guests of the Granthams at Downton. The hit period drama will return to screens this autumn. Series four of the show, which followed the wealthy Grantham family and their servants, achieved an average of 11.9 million viewers in the UK. The very British drama has also been a huge hit in the US, winning both Emmy Awards and Golden Globes. More than 26 million viewers watched series four on Masterpiece on PBS, making it one of the highest rating shows on American television. Previous high profile guest stars include Shirley Maclaine who played Martha Levinson, Lady Grantham's mother, and Oscar-nominated actor Paul Giamatti who appeared in last year's Christmas special as her "maverick, playboy" son. Series five will also feature 24 star Rade Sherbedgia as a Russian refugee who has fled the revolution after World War 1. Earlier this year, executive producer Gareth Neame promised it would have "all the usual highs and lows, romance, drama and comedy".
\\ \hline
Gold & 
The Royle Family actress Sue Johnston is the latest star to join series five of ITV's Downton Abbey in a guest role.
\\ \hline
Model & 
Former Emmerdale actress Sian Johnston is to join the fifth series of Downton Abbey.
\\ \hline
ROUGE2-F1 & 
17.14
\\ \hline
\makecell{Document\ID \#198)} & 
Media playback is not supported on this device Craig Cathcart put the visitors ahead before substitute Simon Church won and scored an 89th-minute penalty. "There were lots of positives out of it even if we'd have come off and lost 1-0. They had a good mentality and attitude," said Coleman. Wales face another Euro 2016 warm-up game against Ukraine in Kiev on Monday. "We look forward to our next challenge now," added Coleman. "The team will change up again, and we'll see how they go again." Striker Church, currently on loan at Scottish Premiership side Aberdeen from Reading, was delighted with his equaliser from the spot. "Northern Ireland were a tough side to play against. They've obviously done well to get where they are and it was a tough game," he said. "We wanted to do well because it was the last time a Wales crowd would see us before the Euros and we wanted to put in a good performance. "I've just got to keep going now and hopefully score some goals. This is a great squad to be part of."
\\ \hline
Gold & 
Wales manager Chris Coleman said he was pleased with his team's performance after they came from behind to draw 1-1 with Northern Ireland in Cardiff.
\\ \hline
Model & 
Wales manager Chris Coleman praised his side's attitude after they came from behind to draw 1-1 with Northern Ireland in Cardiff.
\\ \hline
ROUGEL-F1 & 
80.00
\\ \hline
\makecell{Document\ID \#279)} & 
The early work, carried out on mice and pigs, reveals the protein-infused patch encourages the growth of healthy cells and leads to less scarring. Scarring can be common after a heart attack, making the heart pump less effectively and sometimes fail. Writing in the journal Nature, researchers say the patch may one day revolutionise treatment. During an attack, muscle cells in the heart die because of a lack of blood flow and scientists believe repairing or replacing some of these cells may help reduce long-term damage. In this trial an international team of researchers soaked a collagen patch in a protein known as Fstl1 and stitched it on to the hearts of animals who had experienced heart attacks. Though the protein occurs naturally in healthy hearts, it becomes depleted in a key layer of the heart after an attack. Two weeks later the hearts began to grow fresh muscle cells and new blood vessels, while showing signs of pumping more effectively. Prof Pilar Ruiz-Lozano at Stanford University (which has patented the patch), said: "Many were so sick prior to getting the patch that they would have been candidates for heart transplantation. "The hope is that a similar procedure could eventually be used in human heart attack patients who suffer severe heart damage." Commenting on the study in Nature, Prof Gordana Vunjak-Novakovic at Columbia University, said the work "could lead to entirely new modalities for treating heart infarction". But she cautioned that further studies needed to be done to understand whether this type of approach would work on larger animals and ultimately humans.
\\ \hline
Gold & 
A prototype patch could help the repair the damage caused by a heart attack, scientists say.
\\ \hline
Model & 
Scientists say they have developed a synthetic patch that can repair damaged hearts after a heart attack.
\\ \hline
ROUGEL-F1 & 
36.36
\\ \hline
\makecell{Document\ID \#186)} & 
There could be "serious implications" for Gloucestershire Police, the police and crime commissioner has warned. Cotswold District Council's proposal to form a unity authority with West Oxfordshire has proved controversial. But CDC says the plans - dubbed 'Coxit' - are at an early stage but aim to improve accountability. CDC leader Lynden Stowe has said Gloucestershire's "historic borders" would remain under the proposals, which aim to improve the "administration and the delivery of council services". He told BBC Radio Gloucestershire a few weeks ago: "We would expect the police to patrol up to the existing county borders, exactly as now, so if there's a crime in Cirencester, Gloucestershire Police attend and if there's a burglary in Burford, Thames Valley Police attend." But Gloucestershire's PCC Martin Surl said he had taken legal advice on the "Coxit" proposal, and warned it could "signal the end of Gloucestershire Police as we know it". Under the Local Government and Public Involvement in Health Act 2007, he said the Cotswolds could become part of Thames Valley Police District. "If they break away to a new district, we can't continue policing just half a district," he told the BBC. "So, either we start taking some of Thames Valley's area in Oxfordshire, or they start policing parts of Gloucestershire. Neither want that ." A spokesman for Cotswold District Council said they "fully understand the initial concerns" but they were "best addressed through the detailed studies that will now take place to assess the feasibility of the unitary proposal". Mr Surl, an independent, is standing for re-election as Gloucestershire PCC on 5 May, alongside Labour's Barry Kirby and the Conservatives' Will Windsor Clive.
\\ \hline
Gold & 
The Cotswolds could be policed by the Thames Valley force if plans to "break away" from Gloucestershire County Council go ahead, it is claimed.
\\ \hline
Model & 
Plans to create a new district in the Cotswolds have been criticised by a police chief.
\\ \hline
ROUGEL-F1 & 
15.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on XSum sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{CNN/DailyMail} \\ \hline
\makecell{Document\ID \#256)} & 
kevin de bruyne's agent expects to go 'around the world' discussing his client as interest in the wolfsburg midfielder increases ahead of the summer transfer window. patrick de koster, who has confirmed that he held talks with manchester city earlier this season, admits that he could receive 20 phone calls a day about the belgium international as clubs prepare to strengthen their squads before the start of next season. city are keen to sign the ex-chelsea midfielder as they look to reshape their squad but de koster insists no decision has been made and that the 23-year-old could even remain at the volkswagen arena. wolfsburg midfielder has attracted interest from manchester city, bayern munich and paris saint-germain . de bruyne tussles with schalke's sead kolasinac during wolfsburg's 1-1 draw at the volkswagen arena . speaking to the guardian, de koster said: 'for the moment, there are no formal discussions. of course in the next few weeks i will be going around the world to talk about the situation with kevin but this is just informal information. 'i will talk to everybody but kevin is very, very happy with wolfsburg and the way they have treated him since he arrived from chelsea last year. 'there are still five games to play and hopefully they can make sure of a place in the champions league next season so it is a little bit too early to be making any decisions. he has a four-year contract at wolfsburg so we will have to see what they want to do. 'i have met the people from manchester city and we know each other. i've never spoken to anyone from manchester united but a lot of other clubs have been in touch to find out some general information.' bundesliga champions bayern munich and ligue 1 ... ... ... 
\\ \hline
Gold & 
patrick de koster will go 'around the world' to talk about kevin de bruyne . the wolfsburg midfielder is wanted by manchester city and bayern munich . de koster has admitted having talks with city chiefs this season . but he has not spoken to manchester united about a move for his client . de bruyne remains happy at wolfsburg and could yet remain at the club .
\\ \hline
Model & 
kevin de bruyne has attracted interest from manchester city, bayern munich and psg . patrick de koster expects to go 'around the world' discussing his client . de bruyne has scored 10 league goals and provided 17 assists this season .
\\ \hline
ROUGE1-F1 & 
49.50
\\ \hline
\makecell{Document\ID \#197)} & 
i yield to no one in my love of the old days — warm beer, cricket on the village green, bobbies on bicycles two by two, all that — but it's rare a chance arises to compare the rose-tinted past with the brave new world, as it did on saturday evening when sky's high-octane premier league coverage went head-to-head with arsenal v reading in the fa cup semi-final on the bbc. as we know, the premier league has the money and prestige, but what the fa cup has is history, and boy does the bbc love a bit of history? lest you were in any doubt, its coverage of the semi-final kicked off with footage of the late sir laurence olivier doing the st crispin's day speech from the film of henry v ('we happy few, we band of brothers,' and so on). gary lineker, alan shearer, jason roberts and ian wright fronted the bbc's coverage at wembley . bbc presenter lineker prepares to present the match of the day 50th anniversary special broadcast . reading defender nathaniel chaloboah (left) chases arsenal midfielder aaron ramsey (right) on saturday . gunners forward alexis sanchez celebrates after scoring his side's winning goal in the fa cup semi-final . stand-in match of the day presenter gabby logan (left) with pundits phil neville and robbie savage (right) the excuse, i guess, was that reading's nickname is the royals, but as the second-tier team are also known as the biscuitmen. the bbc even gave us a moment of history at half-time with a breakdown, something that used to be a regular feature of tv outside broadcasts, when a feature on crystal palace's shock semi-final victory over liverpool 25 years ago came to a shuddering halt halfway through. in such instances, the bbc is lucky to ... ... ... 
\\ \hline
Gold & 
the weekend saw bbc's fa cup coverage compete with sky's premier league . it was a refreshing throwback to see the bbc's use of archive footage . gary lineker remains one of the bbc's prized assets and they must keep him .
\\ \hline
Model & 
arsenal beat reading 1-0 in the fa cup semi-final at wembley on saturday . the bbc presented the match of the day 50th anniversary special . the price of live premier league football is now way beyond the reach of the bbc, which may be the clincher in gary lineker's future .
\\ \hline
ROUGE1-F1 & 
29.47
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on CNN/DailyMail sampled by ROUGE1-F1.}
\end{table*} \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{CNN/DailyMail} \\ \hline
\makecell{Document\ID \#148)} & 
a former lager lout who ballooned to 24 stone has lost nearly half his body weight by giving up his favourite drink. rugby prop dale forrest, 26, of bolton, would sink up to 12 pints a night, but decided to shed the pounds after seeing a photo of himself while out with his fitness fanatic friends. in december 2013, the bank teller decided to ditch the booze, give up his favourite fatty readymeals and greasy takeaways and hit the gym. dale forrest would drink up to 12 pints a night - and ballooned to 24 stone - before giving up beer to lose weight . dale, pictured at his slimmest, was worried he would look like a 'beached whale' next to his friends on holiday . since then, mr forrest, who had a holiday planned with his mates and didn’t want to look like a ‘beached whale’ next to them, has lost 10.5 stone. he said: ‘i can now go on nights out and feel confident - even without the dutch courage. ‘i no longer feel like people are laughing at me and all it took was a bit of willpower.’ mr forrest struggled with his weight from a young age and in adulthood fell into unhealthy habits. before shedding the pounds mr forrest ate a cheese and sausage bap for breakfast, a big daddy box meal from kfc for lunch and a meat feast pizza for dinner. dale, pictured sitting down, said he would consume beer, alcopops and shots regularly on nights out . dale, pictured before his weight loss (right) and after losing 10 stone, gave up greasy takeaways for healthy foods and started going to the gym regularly, and saw the pounds fall off . despite being a rugby player, dale weighed 24 stone due to his ... ... ... 
\\ \hline
Gold & 
dale forrest would go out drinking regularly and would eat fatty food . dined on cheese and sausage bap for breakfast and kfc for lunch . decided to lose weight after seeing photos of him next to slim friends . started going to the gym and eating healthy foods and lost 10 stone .
\\ \hline
Model & 
dale forrest would drink up to 12 pints a night on nights out with friends . 26-year-old from bolton worried he would look like a 'beached whale' on holiday . in december 2013 he decided to lose weight after seeing a photo of himself . ditched the booze and started going to the gym and lost more than 10 stone .
\\ \hline
ROUGE2-F1 & 
26.17
\\ \hline
\makecell{Document\ID \#73)} & 
she's a best-selling singer, actress, beauty buff and one of the world's most stylish stars. and now, rita ora - who is only 24 and has already made the transition from a kosovan-born fledgling singer to one of the globe's most successful stars - has channeled her passion for fashion into a new adidas range. for her latest collaboration with adidas originals, rita has taken the brand's classics and put her own bold spin on them. the result? a collection that is as daring and vibrant as the designer herself. scroll down for video . rita ora has channeled her passion for fashion into a new adidas range, so femail caught up with the global star to find out her influences and plans for the future . speaking to femail about the inspiration behind her new range, which is emblazoned with graphic dragon prints and lands on may 1, rita said: 'the dragon print is all about my love of travel and soaking up different cultures. the print itself is inspired by asian culture and one of its most recognisable symbols. 'i’ve contrasted that with the fit, basketball style cuts, which were inspired by american culture. my white smoke pack is about the body-mind-soul connection, about taking a moment for reflection. the graphic comes from the visual smoke creates when burning incense.' the talented star, who also recently unveiled a beauty range for rimmel, couldn't be happier with the final result. she said: 'i love being able to go into design meetings with all these crazy ideas and then be able to create a product that hasn’t existed before. from the initial stages to seeing my fans wearing the clothes and shoes - it’s been an amazing experience.' the talented star, who also recently unveiled a beauty range for rimmel, ... ... ... 
\\ \hline
Gold & 
rita, 24, has designed range for adidas originals . designs are inspired by asian culture, she tells femail . star says she's excited to see what the future holds for her .
\\ \hline
Model & 
rita, 24, has teamed up with adidas originals . has taken brand's classics and put her own bold spin on them . promises there's 'a little something for everyone' in the range . cites marilyn monroe as her ultimate postergirl .
\\ \hline
ROUGE2-F1 & 
8.96
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on CNN/DailyMail sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{CNN/DailyMail} \\ \hline
\makecell{Document\ID \#139)} & 
a father whose 20-year-old daughter was found murdered in iowa last year made a desperate plea from china for u.s. authorities to do more in tracking down her killer, six months after the girl's body was discovered. and police in iowa seem to have now answered it, by issuing a warrant for the girl's boyfriend, according to reports. tong shao, a chemical engineering student at iowa state university, went missing in september 2014. after a three week search, police found her body stuffed in the trunk of her toyota camry in iowa city. shao's boyfriend, xiangnan li, 23, was listed as a person of interest in the case and is believed to have been the last person to see her alive, however he bought a one-way ticket to china in the days after his girlfriend went missing and has disappeared, cnn reported. wanted for murder: an arrest warrant has reportedly been issued for , xiangnan li, 23 (right), the boyfriend of tong shaom 20 (left), a university of iowa student found murdered in september after going missing . li had transferred to iowa from rochester institute of technology to be closer to tong. the two had meet studying english in beijing in 2011. they had checked into a hotel room together on september 5, 2014. the pair had stayed at the same hotel three times before and the owner knew them. according to police records obtained by cnn, two days earlier tong had accidentally called li - or 'pocket dialed' him - and he stayed on the line for 30 minutes, overhearing a conversation. tong was complaining about li to a friend and said things that 'were not nice', the records noted. the owner of the hotel told investigators li left the hotel either on the night of september 6 ... ... ... 
\\ \hline
Gold & 
tong shao, 20, was an international student from china attending iowa state university . her body was found in the trunk of her car in iowa city on september 26 . police believe it had been for three weeks . she died of blunt force trauma and asphyxiation . her boyfriend, xiangnan li, 23, was the last to see her, but flew to china on september 8, before shao was officially missing . according to tong's father, an arrest warrant has now been issued . however li has disappeared .
\\ \hline
Model & 
tong shao, 20, was found murdered in the trunk of her car in september . her boyfriend, xiangnan li, 23, was a person of interest in the case . li bought a one-way ticket to china in the days after tong went missing . he is believed to have been the last person to see her alive .
\\ \hline
ROUGEL-F1 & 
35.97
\\ \hline
\makecell{Document\ID \#55)} & 
danny willett gave a rules official, who had been in his line-of-sight, a verbal blast which was clearly audible to spectators surrounding the 17th green at the masters on thursday. englishman willett vented his anger after his second shot from beyond the green trickled all the way across the putting surface and left the preacher's son facing a bogey or worse. 'of anyone you should know the rules,' willett shouted loudly at the official, who was sitting about 60 yards away in a golf cart, outside the gallery ropes. danny willett waits to play a shot on the fifth during the first round at 2015 masters on thursday . the 27-year-old englishman carded a one-under 71 during his first ever round at augusta . willett was still visibly angry as he left the green after salvaging a bogey, though he had calmed down by time he spoke to reporters about 30 minutes later. he said his ire had been raised because the official had been in his line-of-sight as he was preparing to play his shot. 'we were being timed (for slow play), which i can appreciate,' willett said after carding a one-under 71 at augusta national. 'it's a little bit tricky out here, so it takes a little bit of time, but you'd like to think the referee that's timing you knows exactly where to put his buggy and where not to put his buggy.' willett, 27, a two-time european tour winner, is playing in his first masters. willett was left frustrated with a rules official for getting in his line-of-sight during his round .
\\ \hline
Gold & 
englishman danny willett blasts timing referee for getting in line-of-sight . vented anger at official as he bogeyed the 17th at 2015 masters . willett carded a one-under 71 on opening round in first time at augusta .
\\ \hline
Model & 
danny willett gave a rules official a verbal blast during the first round . the englishman vented his anger after his second shot from beyond the green trickled all the way across the putting surface and left the preacher's son facing a bogey or worse . willett was still visibly angry as he left the green after salvaging a bogey, though he had calmed down by time he spoke to reporters about 30 minutes later . he said his ire had been raised because the official had been in his line-of-sight as he was preparing to play his shot .
\\ \hline
ROUGEL-F1 & 
13.14
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on CNN/DailyMail sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{NEWSROOM} \\ \hline
\makecell{Document\ID \#228)} & 
Seven species of bees in Hawaii have been classified as endangered, the first time the insect has been protected by federal law. The U.S. Fish \& Wildlife Service has granted seven species of yellow-faced bees native to the islands protection under the Endangered Species Act, which will hopefully “allow authorities to implement recovery programs, access funding and limit their harm from outside sources,” Gregory Koob of USFW told The Associated Press. The yellow-faced bee’s population faces a range of threats, like habitat destruction, invasive species both animal and vegetable, and all manner of natural phenomena such as wildfires, hurricanes, tsunamis and droughts. This news represents an important step for bees in general. They — and many other pollinating species like butterflies — are in danger across the world: About 40 percent of invertebrate pollinating species are facing extinction, according to a U.N. report released in February. This could have devastating effects on global agriculture, as about 75 percent of the world’s crops rely on pollination to grow. The ruling also grants protection to 39 plant species found in Hawaii and three other native animals: the band-romped storm-petrel, the orangeblack Hawaiian damselfly and the anchialine pool shrimp. Bees in certain parts of the U.S. are facing a different threat, as aerial spraying, aimed to combat Zika-spreading mosquitos, is killing hives en masse.
\\ \hline
Gold & 
A species of bee in Hawaii has been added to the endangered species list, a first for the insect.
\\ \hline
Model & 
Seven species of bees in Hawaii have been classified as endangered, the first time the insect has been protected by federal law. The U.S. Fish \& Wildlife Service has granted seven species of yellow-faced bees native to the islands protection under the Endangered Species Act, which will hopefully “allow authorities to implement recovery programs, access funding and limit their harm from outside sources,” Gregory Koob of USFW told The Associated Press.
\\ \hline
ROUGE1-F1 & 
28.57
\\ \hline
\makecell{Document\ID \#98)} & 
You may have heard Skip Bayless is a Dallas Cowboys fan. And unlike in past years, that is worth celebrating (which Bayless does quite frequently on Undisputed). At 12-2 the Cowboys are running away from the rest of the NFC, have clinched a playoff berth and have a rookie duo in Dak Prescott and Ezekiel Elliott that appears poised to make a deep playoff push in 2017 and beyond. With such a bright future, many Cowboys fans will be searching for a way to suitably express their jubilation. Enter Bayless and DJ Steve Porter with How Bout Them Cowboys. There are a few observations to be made right off the bat. 1) What is that table on set made of and where can it be ordered? The abuse it has taken without breaking is remarkable. 2) There are an extraordinary number of famous people taken aback by Skip’s passion in 90 seconds. At a glance ... “I have a feeling I’m about to unleash, you’re Cowboy hating, is it Dak is it Zeke?” “My Cowboys have arrived, are you sleep deprived? You have to eat humble pie” “Monday after Monday, they just keep making plays, here we go, every dog has its day” “Dak attack, the MVP, Dak attack, you decree, ah ah ah, kee kee kee, week after week, I do agree” “Romo need to check that team chemistry, Dez Bryant chasing greatness oh that’s fancy” 3) A special consideration should be made by the Recording Academy to bestow an honorary Grammy to the whites of Shannon Sharpe’s eyes for their performance throughout the video. Is it a hit single? Only time will tell. Will the Cowboys use this as inspiration to reach their first Super Bowl since 1996? If they do, someone better reinforce that table.
\\ \hline
Gold & 
It won't end the QB controversy, but it is catchy.
\\ \hline
Model & 
Skip Bayless is a Dallas Cowboys fan. And unlike in past years, that is worth celebrating (which Bayless does quite frequently on Undisputed). At 12-2 the Cowboys are running away from the rest of the NFC, have clinched a playoff berth and have a rookie duo in Dak Prescott and Ezekiel Elliott that appears poised to make a deep playoff push in 2017 and beyond. With such a bright future, many Cowboys fans will be searching for a way to suitably express their jubilation. Enter Bayless and DJ Steve Porter with How Bout Them Cowboys. There are a few observations to be made right off the bat.
\\ \hline
ROUGE1-F1 & 
3.36
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on NEWSROOM sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{NEWSROOM} \\ \hline
\makecell{Document\ID \#32)} & 
TESTED 2013 Lincoln MKZ Hybrid WHAT IS IT? A Lincoln-branded upscale hybrid sedan with its Ford roots showing. HOW MUCH? \42,415 as tested. The price is the same as an equivalent nonhybrid MKZ. WHAT MAKES IT RUN? A 141-horsepower 2-liter 4-cylinder, with an 88-kilowatt electric motor and a 1.4-kilowatt-hour lithium-ion battery for 188 net horsepower; continuously variable transmission. IS IT THIRSTY? The E.P.A. rating is 45 m.p.g. all around    in the city, on the highway and combined. ALTERNATIVES Lexus ES 300h, Ford Fusion Hybrid. Part of the fun of a hybrid car is listening for the gasoline engine to start and analyzing how smoothly it makes the transition from electric to gas. Unless you floor the accelerator, the MKZ Hybrid is so quiet it  s hard to tell what  s moving you forward. With active noise control, which sends sound-canceling frequencies through the infotainment system, the car errs on the side of isolation: if engagement with the road is your goal, this isn  t your ride. Rather, the appeal is the combination of an impressive mileage rating and an understated wood-and-leather ambience. Ford calls the MKZ Hybrid   America  s most fuel-efficient luxury sedan.   Consumers are showing an active interest in upscale gas sippers, a category that never existed in the past. This is the second iteration of the MKZ Hybrid, which shares a powertrain with the Ford Fusion Hybrid. Both cars were extensively reworked for 2013, with striking    but quite different    new designs. The MKZ  s extroverted styling is one of the best things about it. The winged grille treatment recalls a subtler form on the streamlined Stout Scarab of the 1930s. The lithium-ion battery replaces the nickel-metal-hydride pack in ... ... ... 
\\ \hline
Gold & 
The appeal of the 2013 Lincoln MKZ Hybrid is its impressive 45 miles per gallon rating and understated wood-and-leather ambience.
\\ \hline
Model & 
The appeal of the Lincoln MKZ Hybrid is the combination of an impressive mileage rating and an understated wood-and-leather ambience.
\\ \hline
ROUGE2-F1 & 
52.38
\\ \hline
\makecell{Document\ID \#163)} & 
The age when North American clubs look to England and the wider British Isles with a child’s embrace, beseeching input from the mother country’s coaching bosom to take the domestic game on, has long since passed. Or so some critics would argue. A new, forward-thinking generation of coaches not steeped in the 4-4-2 and an unadorned, direct style of football rule the waves in this epoch, they postulate. And this legion of coaches are increasingly young, fresh and, most importantly, American. While there might be some historical merit to the spirit of the argument against the British-style coach, it is perhaps itself a little outdated, not to say somewhat harsh on someone like Carl Robinson and his stylish Vancouver Whitecaps, for instance. Some 20 years on from the birth of Major League Soccer, some of the men who as players helped mold the country’s top division into what it has become today now form an integral portion of the younger cohort of the domestic coaching ranks. Broadly successful elder statesmen like Bruce Arena and Sigi Schmid are followed by a seemingly blossoming rank and file. Head coaches such as Sporting Kansas City’s Peter Vermes. DC United’s Ben Olsen. New England Revolution’s Jay Heaps. Columbus Crew’s Gregg Berhalter. New York Red Bulls’ Jesse Marsch. Behind them, too, are some well-thought-of assistants. Then there are the firmly rooted foreign coaches partly schooled in MLS ways after successful stints in the league. Robinson would be one. As would FC Dallas head coach Oscar Pareja. Related: Would a pan-American Champions League benefit soccer in the US? Yet the country has yet to see a coach leave home shores and make a name with instant recognition. Amid the chattering classes of the global game’s chief draws in Europe, that is not always easy. It is ... ... ... 
\\ \hline
Gold & 
Bryan Kay: Plenty of Americans play abroad, but far fewer have made careers as coaches of top sides outside MLS. Could former USA coach Bob Bradley break that pattern?
\\ \hline
Model & 
American players are now a common presence in the Old World top divisions of England and Germany, but no American manager has yet made a name for himself
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on NEWSROOM sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{NEWSROOM} \\ \hline
\makecell{Document\ID \#127)} & 
Poverty comes in many forms: a middle-aged man making \10 an hour, a single mother with mouths to feed, a person with a criminal record and few employers willing to look past it.
\\ \hline
ROUGEL-F1 & 
23.53
\\ \hline
\makecell{Document\ID \#95)} & 
“Y'all think a little toy buzzer is going to keep the truth from tumbling out from these jaws?” With news that Olivia's (other) ex Edison Davis (Norm Lewis) is joining the race, Shondaland has completed the fantasy: two women, a Hispanic man, a black man and a racist redneck who is almost certainly but not actually Donald Trump. 'SCANDAL' RECAP SEASON 5, EPISODE 14: BILLIONAIRE BUFFOON RUNS FOR PRESIDENT — BUT NOT THE ONE YOU'RE THINKING OF Yes somehow, this election still isn't interesting. Cyrus (Jeff Perry) is fighting for control of Francisco Vargas (Ricardo Chavira) with the candidate's brother, which is mind-numbingly boring. Olivia (Kerry Washington) and Huck (Guillermo Diaz) have set up a focus group on Mellie's (Bellamy Young) campaign, which lends itself to a hilariously ridiculous new slogan: “Mellie Grant, woman of the people.” Remember when she got drunk on hooch and ate fried chicken for months because a secret society killed her son? Quinn (Katie Lowes) goes undercover as a WASP to gather information on Vanessa, because Liv's still freaking out about Jake's (Scott Foley) love life. And Hollis Doyle is shooting really big guns and posing for cameras. Election 2016, y'all. It's weird in all realities. Mellie's trek to Gettysburger to relate to the commonfolk backfires when she tells reporters that she and Karen (remember her?) used to go every Sunday after church. But — dun-dun-dun! — Gettysburger is closed on Sundays. So we have Burgergate. End the meme. End it now. 'SCANDAL' RECAP 5X12: FITZ IS DATING, CYRUS IS SCHEMING, OLIVIA IS SPYING After some more WASP undercover work by Quinn, we finally get some marginally interesting information: Jake is stealing money from Vanessa to give to Papa Pope (Joe Morton) to fund a Super Pac for Edison. Which, at least somewhat, explains why ... ... ... 
\\ \hline
Gold & 
With news that Olivia's (other) ex Edison Davis (Norm Lewis) is joining the race, Shondaland has completed the fantasy.
\\ \hline
Model & 
“Y'all think a little toy buzzer is going to keep the truth from tumbling out from these jaws?”
\\ \hline
ROUGEL-F1 & 
10.26
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on NEWSROOM sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Multi-News} \\ \hline
\makecell{Document\ID \#145)} & 
Image copyright Crossrail Image caption The plague victims' bones reveal clues to their harsh lives in medieval London Skeletons unearthed in London Crossrail excavations are Black Death victims from the great pandemic of the 14th Century, forensic tests indicate. Their teeth contain DNA from the plague bacterium Yersinia pestis and their graves have been dated to 1348-50. Records say thousands of Londoners perished and their corpses were dumped in a mass grave outside the City, but its exact location was a mystery. This discovery solves a 660-year-old mystery. It's a hugely important step forward Jay Carver , Lead archaeologist, Crossrail Archaeologists now believe it is under Charterhouse Square near the Barbican. They plan to expand their search for victims across the square - guided by underground radar scans, which have picked up signs of many more graves. Crossrail's lead archaeologist Jay Carver says the find "solves a 660-year-old mystery". "This discovery is a hugely important step forward in documenting and understanding Europe's most devastating pandemic," he said. "Further excavations will follow to see if - as we expect - we are coming across a much bigger mass burial trench." Media playback is unsupported on your device Media caption The teeth of the skeletons contain plague bacterial DNA Between 1347 and 1351 the "Great Pestilence" swept westward across Europe killing millions of people. It later became known as the Black Death. The plague The plague is one of the oldest identifiable diseases known to man Plague is spread from one rodent to another by fleas, and to humans either by the bite of infected fleas or when handling infected hosts Recent outbreaks have shown that plague may reappear in areas that have long been free of the disease Plague can be treated with antibiotics such as streptomycin and tetracycline Source: World ... ... ... 
\\ \hline
Gold & 
– Skeletons dug up in London last year are indeed the remains of people who died from the Black Plague—and who suffered a tough life before falling ill, the BBC reports. Forensic analysis shows that teeth taken from at least four of the 12 corpses discovered during excavation for a rail line contained trace amounts of plague DNA, indicating exposure. Early burials found at the site, from the late 1340s, are nice and orderly, with bodies wrapped in white shrouds, but skeletons from a second outbreak in the 1430s are tossed in with what appear to be upper-body injuries—evidence of "a period of lawlessness and social breakdown," Phys.org reports. Among other significant details: Several skeletons suffered from malnutrition and 16\% had rickets. Many had back damage, signalling stressful manual labor. Analysis of one victim is amazingly detailed: He was born outside of London, breastfed, had bad tooth decay as a boy, and worked in manual labor before dying as a young man of the Black Death. Archaeologists suspect that thousands more Black Death victims lie nearby, and a dig is planned for this summer. DNA experts are analyzing the plague genome in victims' teeth in case there's more to learn about the disease, which still kills 2,000 people per year (
\\ \hline
Model & 
– It's a "hugely important" find, one that could shed light on the Black Death, the plague that wiped out at least 60\% of London's population between 1347 and 1351, the BBC reports. Archaeologists digging for railway tracks in London's Crossrail station uncovered 25 skeletons, along with pottery dated to the mid-14th century, that have DNA from the bacterium Yerinia pestis dated to 13 to 50 years old. That's a long way from the 660-year-old record showing that thousands of Londoners died from the plague and their bodies were dumped in a mass grave outside the City of London. "We can start to answer questions like: where did they come from and what were their lives about?" says Crossrail archaeologist Jay Carver. "I'm amazed of the reasons for why the Black Death was so devastating." DNA from 12 of the skeletons has confirmed that they had contact with the plague prior to their death, and researchers plan more excavations to see if they're coming across a much bigger plague trench. "We can see that Londoners weren't living an easy life," Carver says, "and that's possibly one of the explanations for why the Black Death was so devastating." Antibiotic-resistant strains of the plague
\\ \hline
ROUGE1-F1 & 
34.58
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Multi-News sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Multi-News} \\ \hline
\makecell{Document\ID \#176)} & 
Sandra Chesterman, left, puts a braclet on the wrist of Christine Kirby during a birthday party for her son Jayden Kirby, 1, on Sunday, Feb. 9, 2014, in Fremont, Calif. Chesterman's daughter Kristina was killed by an alleged drunk driver last September while coming home from nursing school in Chico. Jayden was given months to live before he received part of Kristina Chesterman's liver. Jayden and Christina share the same birth stone and that gem is embeded into the bracelet. (Aric Crabb/Bay Area News Group) ( ARIC CRABB ) LIVERMORE -- When she was still in high school, Kristina Chesterman wrote out her bucket list. Flying a plane was on it; so was running through a poppy field and breaking up a fight between two boys over her affections. She also wanted to save a life. The aspiring nurse's ambitions came to a halt in September, when she was killed by a suspected drunken driver near Chico State, where she attended school. Though Chesterman, 21, didn't get to mark much off her list, she has saved more lives than she hoped -- and is profoundly affecting many others. Five Northern Californians have been saved through Kristina's choice to donate her organs. And her grateful friends and family are making the rest of her bucket list their own. A photograph of Kristina Chesterman is displayed during a birthday party for Jayden Kirby on Sunday, Feb. 9, 2014, in Fremont, Calif. Chesterman was killed by an alleged drunk driver last September while coming home from nursing school in Chico. Jayden, 1, was given months to live before he received part of Kristina Chesterman's liver. (Courtesy of the Chesterman Family) ( Chesterman Family ) Chesterman's mother, Sandra, of Livermore, said her daughter wanted to help people from an early age. She routinely gave ... ... ... 
\\ \hline
Gold & 
– Kristina Chesterman, 21, was studying to be a nurse when she was killed by a suspected drunk driver last year—but she managed to save lives anyway. A registered organ donor, Chesterman gave five people, including a baby, new life, and now the woman who received her heart wants to do something in return. Susan Vieira, 64, has vowed to check off everything on Chesterman's bucket list—written down on a piece of paper her mom only recently found. (One of the items? "Save someone's life." Another? "Be in four places at once.") Vieira had completed several of the tasks already, including learning to fly a plane and riding a camel, ABC News reports. And now, "together, we will finish her bucket list," Vieira says. Other points on the list include running through a poppy field and riding in a hot-air balloon, the San Jose Mercury News reported earlier this year. "I'd like to think all the things I continue to accomplish in my life, I’m taking Kristina with me," Vieira continues; Chesterman's mom adds that she "felt an instant connection" to the woman she just met. Chesterman's friends are also helping to complete the bucket list, including
\\ \hline
Model & 
– When 21-year-old Kristina Chesterman was killed by a suspected drunk driver in September, she left behind a bucket list of things she wanted to do before she died. Among them: fly a plane, break up a fight between two boys over her affections, and donate her organs. Now, five Northern Californians have been saved thanks to her decision, the San Jose Mercury News reports. "I know she would've been so proud," says Chesterman's mother, Sandra. "It hasn't been an easy process for us, but at the same time, it's brought us so much comfort." One of those saved was 1-year-old Jayden Kirby, who might have died as an infant had he not received part of Chesterman's liver. Jayden's mother, Christine, figured out the name of the mystery donor after hearing news reports, the Mercury News reports. "I wanted her to see the impact Kristina had," she says. "(The transplant) absolutely saved (Jayden's) life." Chesterman's aunt, Patricia Picard, texted the couple, asking if Zak Pappachan was a match. Within hours, he had his new kidney. Now he's enjoying being a dad to his young daughter, Ray
\\ \hline
ROUGE2-F1 & 
10.97
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Multi-News sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Multi-News} \\ \hline
\makecell{Document\ID \#87)} & 
The text on the video above was edited on June 1, 2018 to make it clear that it is impossible to know why the polar bear pictured was starving. An earlier version of the video went too far in suggesting that climate change was responsible ( read more ). This story was updated on January 19, 2018 to reflect the more specific location of where the photographs were taken. When photographer Paul Nicklen and filmmakers from conservation group Sea Legacy arrived on Somerset Island—near the larger Baffin Island—in the Canadian Arctic in late summer, they came across a heartbreaking sight: a starving polar bear on its deathbed. Nicklen is no stranger to bears. From the time he was a child growing up in Canada's far north the biologist turned wildlife photographer has seen over 3,000 bears in the wild. But the emaciated polar bear, featured in videos Nicklen published to social media on December 5, was one of the most gut-wrenching sights he's ever seen. "We stood there crying—filming with tears rolling down our cheeks," he said. Video shows the polar bear clinging to life, its white hair limply covering its thin, bony frame. One of the bear's back legs drags behind it as it walks, likely due to muscle atrophy. Looking for food, the polar bear slowly rummages through a nearby trashcan used seasonally by Inuit fishers. It finds nothing and resignedly collapses back down onto the ground. In the days since Nicklen posted the footage, he's been asked why he didn’t intervene. "Of course, that crossed my mind," said Nicklen. "But it's not like I walk around with a tranquilizer gun or 400 pounds of seal meat." And even if he did, said Nicklen, he only would have been prolonging the bear's misery. Plus, feeding wild polar bears ... ... ... 
\\ \hline
Gold & 
– An estimated 2.5 billion people saw the image: a starving polar bear struggling across an Arctic landscape. "The mission was a success, but there was a problem: We had lost control of the narrative," writes Cristina Mittermeier in National Geographic. Accompanied by a photographic team, she snapped shots of the dying bear last year while colleague Paul Nicklen shot video—all part of their "mission to capture images that communicate the urgency of climate change," she writes. "When Paul posted the video on Instagram, he wrote, 'This is what starvation looks like.'" He also wondered if all 25,000 polar bears would die like this and urged people to reduce their carbon footprint. But he didn't say climate change had killed this bear. That nuance vanished when National Geographic posted the video with the subtitles "This is what climate change looks like" and the Washington Post ran a dramatic headline about "gut-wrenching" images. People also responded dramatically, expressing gratitude for validation of climate science, anger the crew hadn't fed the bear, or a stubborn refusal to acknowledge global warming. "Perhaps we made a mistake in not telling the full story—that we were looking for a picture that foretold the future and that we didn’t know what had happened to this particular polar bear
\\ \hline
Model & 
– "We stood there crying—filming with tears rolling down our cheeks." That's Paul Nicklen's take on a video he posted in December showing a starving polar bear on an island in the Canadian Arctic, National Geographic reports. The video was shot by Nicklen, a biologist and contributor to National Geographic, during a trip with the conservation group Sea Legacy. It shows the bear on Somerset Island, near the larger Baffin Island, as it searches a garbage can for food. At one point, the bear shows signs of starvation. "When scientists say bears are going extinct, I want people to realize what it looks like," Nicklen tells National Geographic. "Bears are going to starve to death. This is what a starving bear looks like." The video has since been viewed more than 3.5 million times, and Nicklen says it struck a nerve. "If you have any conscience at all, this video of a starving polar bear will break your heart," Newsweek states.
\\ \hline
ROUGEL-F1 & 
17.23
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Multi-News sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Gigaword} \\ \hline
\makecell{Document\ID \#208)} & 
a zairean diplomat was among three african men arrested for smuggling \#\# kilograms -lrb- \#\#\# pounds -rrb- of marijuana into sweden , the swedish news agency tt reported friday .
\\ \hline
Gold & 
zairean diplomat among three arrested for pot
\\ \hline
Model & 
zairean diplomat arrested for drug smuggling
\\ \hline
ROUGE1-F1 & 
61.54
\\ \hline
\makecell{Document\ID \#92)} & 
reinsurer hannover re ag said thursday that its second-quarter net profit doubled as demand increased for its products , particularly life and health reinsurance .
\\ \hline
Gold & 
hannover re s \#nd-quarter profit doubles
\\ \hline
Model & 
hannover re says \#q net profit doubled
\\ \hline
ROUGE1-F1 & 
42.86
\\ \hline
\makecell{Document\ID \#149)} & 
los angeles might have lost its bid for an expansion team to houston on wednesday , but fans who like to watch football and the tv networks were winners .
\\ \hline
Gold & 
no l.a. team good for tv viewers
\\ \hline
Model & 
los angeles loses bid for nfl team
\\ \hline
ROUGE1-F1 & 
26.67
\\ \hline
\makecell{Document\ID \#258)} & 
the suicide bomb attacks in saudi arabia were    a cowardly and disgraceful terrorist atrocity , '' prime minister tony blair said wednesday .
\\ \hline
Gold & 
two britons missing after saudi suicide blasts
\\ \hline
Model & 
blair calls saudi attacks cowardly and disgraceful terrorist atrocity
\\ \hline
ROUGE1-F1 & 
12.50
\\ \hline
\makecell{Document\ID \#39)} & 
chelsea owner roman abramovich and england captain david beckham top english soccer 's rich lists .
\\ \hline
Gold & 
second test scoreboard
\\ \hline
Model & 
abramovich beckham top rich list
\\ \hline
ROUGE1-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Gigaword sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Gigaword} \\ \hline
\makecell{Document\ID \#57)} & 
greece international central defender sotiris kyrgiakos says he intends to join liverpool after receiving an offer from the premier league club .
\\ \hline
Gold & 
kyrgiakos says intends to join liverpool
\\ \hline
Model & 
greece defender kyrgiakos to join liverpool
\\ \hline
ROUGE2-F1 & 
40.00
\\ \hline
\makecell{Document\ID \#13)} & 
the weather was cool during georgia tech 's football practice monday , but coach george o'leary was hot and sweaty after the workout .
\\ \hline
Gold & 
o'leary gets defensive after loss to bc
\\ \hline
Model & 
o'leary feels the heat
\\ \hline
ROUGE2-F1 & 
18.18
\\ \hline
\makecell{Document\ID \#291)} & 
nearly a year ago , south carolina supreme court chief justice jean UNK said that fair , swift justice has taken a backseat to an assembly line process that diminishes the quality of hearings afforded south carolinians .
\\ \hline
Gold & 
fair swift justice demands diversity
\\ \hline
Model & 
s.c. supreme court chief justice backs off criticism
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\makecell{Document\ID \#173)} & 
marc rosset upset boris becker in straight sets to pace switzerland over germany in the opening round of the world team cup monday , while spain beat sweden \#-\# .
\\ \hline
Gold & 
switzerland spain win opening matches
\\ \hline
Model & 
rosset upsets becker in world team cup
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\makecell{Document\ID \#59)} & 
his leftist fringe party does n't even have \#\#\# members , but \#\#-year-old budiman UNK is accused of being at the heart of a communist plot to topple the government .
\\ \hline
Gold & 
UNK weekly sports calendar
\\ \hline
Model & 
fringe party leader accused of plotting to topple government
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Gigaword sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Gigaword} \\ \hline
\makecell{Document\ID \#248)} & 
wednesday 's evening rubber prices in singapore cents per kilo provided by the singapore commodity exchange : UNK
\\ \hline
Gold & 
evening rubber prices
\\ \hline
Model & 
wednesday 's evening rubber prices
\\ \hline
ROUGEL-F1 & 
75.00
\\ \hline
\makecell{Document\ID \#15)} & 
prime minister of antigua and barbuda baldwin spencer left here monday for hong kong , winding up his four-day visit to shanghai .
\\ \hline
Gold & 
antigua and barbuda pm leaves shanghai for hong kong
\\ \hline
Model & 
antigua and barbuda prime minister ends shanghai visit
\\ \hline
ROUGEL-F1 & 
47.06
\\ \hline
\makecell{Document\ID \#164)} & 
russia 's defense industry has been badly hit by the global financial crisis , deputy prime minister sergei ivanov , a former defense minister , said tuesday .
\\ \hline
Gold & 
russia 's defense sector hit by financial crisis : govt official
\\ \hline
Model & 
financial crisis hits russian defense industry
\\ \hline
ROUGEL-F1 & 
25.00
\\ \hline
\makecell{Document\ID \#265)} & 
us defense secretary donald rumsfeld said thursday the killing of al-qaeda 's leader in iraq , abu musab UNK , was a significant victory in the battle against terrorism but not the end of the violence .
\\ \hline
Gold & 
rumsfeld calls zarqawi death significant victory
\\ \hline
Model & 
rumsfeld hails killing of al-qaeda leader in iraq
\\ \hline
ROUGEL-F1 & 
13.33
\\ \hline
\makecell{Document\ID \#51)} & 
the rope on the flagpole is still broken .
\\ \hline
Gold & 
british reopen their baghdad embassy but they ca n't call it that yet
\\ \hline
Model & 
UNK rope still broken
\\ \hline
ROUGEL-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Gigaword sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{WikiHow} \\ \hline
\makecell{Document\ID \#56)} & 
Lip-plumping products come in many forms: glosses, balms, sticks, gels, and pots. Applying these to your lips can temporarily make your lips look fuller, often by irritating them.The plumping effect will only last for a couple of hours, but you can renew it by reapplying the product to your lips. Note that the effects won’t be as drastic as those achieved through cosmetic surgery.; , Ingredients like cinnamon, ginger mint, wintergreen, and capsicum will boost blood flow to your lips, causing them to redden and swell — i.e., to look fuller., If you wish to combine your lip-plumping product with a lipstick or gloss, apply the plumper to your lips first for maximum effectiveness. Dermatologists recommend against overusing lip-plumping products, as these may cause your lips to become dry and scaly.Try to save the lip plumpers for special occasions. If you want more from your lip-plumping products, consider trying a treatment plumper. Manufacturers claim that treatment plumpers can stimulate your lips into producing more collagen and elastin, making your lips fuller for a longer period of time. Treatment plumpers are available for purchase online and in person at shops that carry beauty products.They may be more expensive than traditional plumpers. Common ingredients in treatment plumpers include peptides, marine collagen, and human growth factors.
\\ \hline
Gold & 
Invest in a lip-plumping product. Know which lip-plumping ingredients to look for. Use a lip plumper as a primer. Don’t overuse lip plumpers. Look into treatment plumpers.
\\ \hline
Model & 
Apply a lip-plumping product. Use natural lip-plumping ingredients. Combine lip-plumping products. Try a treatment plumper.
\\ \hline
ROUGE1-F1 & 
50.00
\\ \hline
\makecell{Document\ID \#3)} & 
If you hang up on a telemarketer right away, you’ll probably be placed on the “no answer” list, and will be called again eventually. You don’t want to engage a telemarketer in conversation either, especially if you have absolutely no interest in their product or service. The easiest way to handle a telemarketer is to say, “Please put me on your do not call list.” If the telemarketer keeps interrupting you or a robot calls you, you might just have to hang up. If the same number persistently calls, report it to the FCC by calling 1-888-CALL-FCC or going to https://www.fcc.gov/. If you live in the United States, you can add yourself to the do not call registry. It'll then become illegal for telemarketers to contact you: https://www.donotcall.gov/. If you do get contacted by a telemarketer after putting yourself on the list, you can report the number that called you to the FTC or FCC. If they tell you a mutual acquaintance or other contact, consider taking the call. If they don’t or can’t tell you, or if they got your contact information from some data pool or list, ask to be removed from their call list and hang up., Be clear about conveying your annoyance with phone calls, especially if you might want to do business with or purchase something from that company but don’t want annoying callers. Make sure you provide them with an email address specifically for spam or sales deals.
\\ \hline
Gold & 
Don’t just hang up right away. Add yourself to the national do not call registry. Ask where they got your contact info. Tell them “No calls, emails only.” If you’re running a business, in charge of purchasing, or otherwise need to field sales pitches from time to time, tell telemarketers that you only respond to emails.
\\ \hline
Model & 
Hang up immediately. Report persistent callers. Ask to be removed from the call list. Be clear about your preferences.
\\ \hline
ROUGE1-F1 & 
20.78
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on WikiHow sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{WikiHow} \\ \hline
\makecell{Document\ID \#193)} & 
While such snacks are not necessarily healthy, they do tend to be lower calorie than other candy options. If you want to buy snacks in the theater, these are generally your best choice of candy. In addition to being lower calorie than other varieties, they are chewier. This means they're more difficult to eat. If you take longer to finish your candy, you're less likely to return to the concession stand for a second helping.Twizzlers have the lowest calorie count, containing only about 165 calories. Milk duds contain 370 calories, while Starbursts contain 480 calories. Keep in mind, if you're on a specific diet plan that restricts sugar, this is not the best route for you. While these snacks are lower calorie, the bulk of the calories come from sugar.; , Many movie theaters now sell protein bars at the counter. You can also sneak one in from a nearby gas station or drug store. Lower calorie protein bars may contain some ingredients you're craving, like nuts and chocolate, while still having some nutrients.Remember to read nutritional labels. Some protein bars may be misleading. In general, it's best to skip bars with a lot of added sugar. Swedish Fish are much lower in sugar than the average movie snack. They also tend to have a tougher texture, making chewing them difficult. If you slow down the pace of your eating, you will eat less overall.A serving of Swedish Fish is about 19 pieces, roughly a large handful, and contains only 140 calories and 29 grams of sugar. Movie theater containers tend to be smaller, so you may find a container that has a single serving. If you can't find a smaller container, try splitting one with a friend. As it can be tricky to find candy that's lower calorie at ... ... ... 
\\ \hline
Gold & 
Go for Twizzlers, Milk Duds, or Starbursts. Buy a protein bar. Look for Swedish Fish. Ask about the theater's food policy.
\\ \hline
Model & 
Buy chewier candy. Try a lower calorie protein bar. Go for Swedish Fish. Find a smaller container. Bring your own snacks.
\\ \hline
ROUGE2-F1 & 
19.51
\\ \hline
\makecell{Document\ID \#272)} & 
Instead of eating a ton of sugar, which is tempting to do when you're tired, try eating complex carbohydrates and protein. Sugar can give you a rush of energy, but then your energy will crash. Complex carbohydrates, on the other hand, offer more sustained energy, especially when you pair them with protein.For instance, try some natural peanut butter on whole-wheat bread or a piece of fruit with a slice of cheese. Dehydration can lead to fatigue. Therefore, staying hydrated will help keep your energy up. Try sipping on water throughout the day to make sure you get enough. The Institute of Medicine recommends that men drink 13 cups and women drink 9 cups of water every day.Other liquids can contribute to your water intake, such as juice and coffee. However, drinking too much juice can pack on extra calories and cause a sugar crash, so try to limit your intake. Also, though you can count caffeinated beverages, you shouldn't make them the majority of the liquid you drink in a day., Keeping something in your mouth, such as a piece of gum, can help increase your alertness. The best time for this type of trick is when you're in a meeting that you absolutely need to stay awake for.
\\ \hline
Gold & 
Skip the sugar. Make hydration a priority. Try a piece of gum.
\\ \hline
Model & 
Eat the right foods. Stay hydrated. Chew something.
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on WikiHow sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{WikiHow} \\ \hline
\makecell{Document\ID \#150)} & 
This area does not need to be perfectly square or rectangular unless you are planning on using sod instead of simply replanting your grass. Drive wooden or plastic stakes around the area to be leveled.; , Attach strings to stretch between the stakes, a few inches above the ground. Attach a string level to the strings to determine the high point. This will usually be your starting point and the rest of the ground will be brought to here, but you can also lower some ground if that makes more sense for your project. Using a tape measure and your level, adjust the strings until you can see how much height needs to be added or taken away across the area you are adjusting. Keep in mind that you may want to adjust the grade of your land to help combat drainage problems. Your land should be graded 1" for every 4', leading away from your home. If you are leveling a small area and it is mostly level, you probably will not need to remove the grass. However, if you have a large area and a lot of leveling to do, removing the grass will be much easier. A simple shovel is all that's necessary for a reasonable amount of space. Depending on how much ground you have to cover and what will be going on the ground afterwards, you'll need to level your land with differing mixtures of soil, sand, and compost/manure fertilizer. If you want to grow grass in this area, the cover will need to be nutrient rich. If you are simply wanting to place a small shed or pool, soil and sand will do just fine. Use a garden rake to spread the material evenly, checking using your level and a measuring tape to make ... ... ... 
\\ \hline
Gold & 
Stake off your area to level. Use a string level. Adjust the strings. Adjust for grade. Remove the grass if necessary. Add your ground cover. Spread the topsoil. Tamp the soil. Let it settle. Spread your seeds. Lightly cover with more soil. Water lightly. Reseed as necessary. Alternatively, buy sod.
\\ \hline
Model & 
Determine the shape of the area to be leveled. Set a string level. Adjust the string level. Decide if you will need to remove the grass. Level the land. Compact the soil. Wait for the soil to settle. Purchase the grass. Plant the grass.
\\ \hline
ROUGEL-F1 & 
29.79
\\ \hline
\makecell{Document\ID \#38)} & 
Read books and learn all you can about dementia so you better understand what you are dealing with. There are a wide range of products available to help make both the dementia patient and the caregivers life easier.; , People with dementia are for the most part caring, smart, funny and always wanting to help others. They respond to kindness and they need a lot of patience. Sometimes people with dementia become anxious or angry. At these times diversion is most important. They like to be kept busy with games, crafts, singing, or just talking about the past. Their abilities will be different because the stage of their dementia is different. It will be hard for a mate not to grieve the loss of his/her mate of many years, or for the child not to grieve the loss of a loving parent. Attend support groups and seminars, and know you are not alone. For those who are caring for relatives suffering with dementia, there are a range of dementia products and dementia aids designed to make both carer and patients' lives easier. Many drug stores and medical supply shops may carry them and you can find sites on the web which carry a whole range of items specifically geared towards patients with dementia. These products cover a range of categories, such as memory aids for dementia, mobility support aids, products to improve bathing safety, assistive devices, nutritional items for specific needs, products to help deal with and manage incontinence, special clothing and aids for vision and hearing. Some of these products are designed to help daily life. Dementia products such as mobility aids and hearing devices provide essential day to day support and good nutritional products can provide valuable calories, nutrients and support a balanced diet, or act as a ... ... ... 
\\ \hline
Gold & 
Look for information. Be kind and patient. Use diversion. Find support. Consider buying dementia aids.
\\ \hline
Model & 
Educate yourself. Be patient and kind. Keep them occupied. Don't forget that people with dementia are different. Get support. Look for dementia products.
\\ \hline
ROUGEL-F1 & 
20.51
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on WikiHow sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Reddit TIFU} \\ \hline
\makecell{Document\ID \#295)} & 
this happened about 3 years ago at my best friend's apartment. i had come over to meet our friend's new boyfriend and go out for a fun night of bar hopping and meeting up with more friends. i wanted to do my makeup and hair since i had showered before coming over and didn't have time prior. now, my best friend keeps her bathroom extremely hot since she has a pet snake and it was the middle of winter...so i started to kinda sweat and remembered that i forgot deodorant and i was two seconds from smelling like a hot dude in a sauna with very poor hygiene. i start panicking and getting even more sweaty when i notice the spray bottle on a small shelf. the outside of it was black and purple, i wrongly assumed it was safe to put on the ol pits and go crazy spraying them nice and good. i felt better, panic attack relieved and i come out of the bathroom. fast forward maybe 30 seconds and i start to feel like my under arm skin is slowly and mercilessly being ripped off. i'm wincing in pain and the three of them are just like, what...the hell? i confess and tell what i did and my best friend just starts dying from laughter, then goes into a fit of concern because i need to get the crap off of my skin asap. i washed it off and had really really horrible chemical burns on my arm pits for about 10 days and couldn't put deodorant on for the duration while my skin healed. needless to say i'm sure everyone loved my natural smell for that period of time... probably one of the more painful (mostly painful) and a little embarrassing moments of my life, ... ... ... 
\\ \hline
Gold & 
i sprayed home fragrance on my pits, burned my skin and smelled gross for over a week/uncomfortable discussion with my family doctor
\\ \hline
Model & 
forgot deodorant in my best friend's bathroom and sprayed it on my arm pits.
\\ \hline
ROUGE1-F1 & 
31.58
\\ \hline
\makecell{Document\ID \#297)} & 
i and a friend i will call tony were meeting two other friends (steven and logan) for dinner before our upcoming gaming session. steven and logan arrive at smash burger about a minute before us. we beheld a sight that will forever be ingrained in my mind. i just wish i could have memorialized it for posterity. it had been snowing that day, and there were 5 inches already on the ground. steven and logan pulled into the parking lot in his truck. logan gets out of the truck but slips and falls on a patch of ice covered under the snow. steven walks around the truck to see where logan disappeared. he was greeted by the site of a baby seal flopping on the ice floor struggling to get up. when tony and i pulled in, steven was dragging logan across the parking lot to a drier patch of ground. he was regaining his dignity just as steven was telling us the story. i just wish i had the sense to capture the moment. but, my mind was still processing what i was seeing. dammit, this almost surpasses steven's own [embarrassing flabalanche of 2009](http://www.reddit.com/r/askreddit/comments/977hp/anyone\_with\_pictures\_or\_video\_of\_a\_400lb\_man/)
\\ \hline
Gold & 
went to meet friends for dinner. was met with the sight of a snow eel being dragged over a patch of ice. laughed too hard to take picture.
\\ \hline
Model & 
friend dragged a baby seal across a parking lot covered in snow.
\\ \hline
ROUGE1-F1 & 
20.00
\\ \hline
\makecell{Document\ID \#293)} & 
so, context. i currently have summer job working in a local amusement park in a small town within northern ireland. one of the requirements for this job is to operate the ghost train carts as they go in one door and come out the other, while allowing customers on. using a operating desk, with three buttons on it, i am able to control when cars go in, when the ones in the middle can be moved up and when the ones that have just come out the bottom (with people in them) can be moved up once the customer has left the cart. so...when the customer comes out the bottom door of the ghost train, they are about a 20ft gap away from due to the space for the other carts to take up. because of this i try to make them leave their cart asap to prevent another coming out and crashing into them, and some times i'll press the button to control that cart just a tad to move it so they get the idea to move out quickly. one day a woman comes out the bottom of the ghost train....and just sits there. it's a busy day, im trying to deal with a lot of customers, and i have 6 other people breathing down my neck trying to get on. i tell them i need to get this woman off before they can go on. so, i hit the bottom and shunt her car while she's only starting to try and sit up, she looks shocked at me. i do it again, she looks even more shocked. i start staring at her, implying "please hurry up. you're taking your time" and she slowly moves her way out of the car, grabbing the bars around her and slowly ... ... ... 
\\ \hline
Gold & 
i tried to get a woman out of her ghost train cart quickly by shunting it, i forgot she was disabled.
\\ \hline
Model & 
i shunted a woman with cerebral palsy out of a ghost train.
\\ \hline
ROUGE2-F1 & 
19.35
\\ \hline
\makecell{Document\ID \#251)} & 
so a little background first: i'm on my school's student council, and today was our junior prom assembly. we had a lot going on, much of it relying on a fast internet connection to work (live cameras streamed over appletv, etc.) we realized that as everyone settled into the auditorium, we'd have a large portion of the student body connecting to the school wifi, and we were worried about slowdown. our live camera was basically useless without a good internet connection, so i had the bright idea to hook up the appletv to my personal hotspot. it connected automatically, i ran the feed for a few seconds, worked like a charm, so we moved on. assembly starts, about 15 minutes in and everything is going great. i'm backstage when i see our principal run up to microphone and tell everyone to exit the school immediately, as there is a potential terror threat in the area (i live in an area where things like this never happen; we've had one lockdown in my 10 years of schooling here) obviously, everyone's more than a little worried. everyone files out, myself included. as i'm standing outside, a friend walks up and tells us what he found out from a teacher. one of the students, while trying to connect to the wifi, had "discovered" a network... my network: "taliban secure communications." needless to say, the student went to a teacher, things escalated, and the school ordered a full evacuation. i talked to the principal (a very, very awkward conversation) and got everything sorted out. everyone's parents are panicking, checking their kids out of school now. my personal hotspot is now called "i love school."
\\ \hline
Gold & 
turned on personal hotspot with the name "taliban secure communications" for an assembly, school got evacuated because of a supposed terror threat.
\\ \hline
Model & 
i connected my school's live camera to my personal hotspot, one of the students found out and evacuated the school.
\\ \hline
ROUGE2-F1 & 
4.88
\\ \hline
\makecell{Document\ID \#29)} & 
this started two days ago and it followed up today. so we had a new sales guy start at the office. i handle most of the i.t. helpdesk/network admin stuff as well as work in sales (its a private company). i was training him on our systems on his first day and it got to around noon and a little bit past so we went out to lunch. chipotle! we're talking about stuff getting to know each other. people always tend to ask questions about college, family blah blah blah. he asks me if i had any siblings, now this question is always tricky. well, my sister died 7 years ago, and i'm a pretty chill guy so i don't like to bring it up and be the debby downer so i usually do a little white lie. if i have no connection whatsoever back to my real life i say 1, if not i say 2. so i said i have 2, i **was** the middle child growing up, (haven't lied yet). hbu? he goes on saying he has an older sister.. nice dodged a bullet. so i make sure it went off on his topic so i didn't have to lie through my teeth. i hate people feeling sorry for me, compassion isn't really a trait that i have (before or after). so fast forward to today when he asks me to go to lunch again today. i try to keep my distance with people who i just meet just so i don't seem like a stage 5 clinger. so i tell him the good place around the area and we go. this is when it gets bad.. he brings back up, "so what do your sisters do" well i say my little sister is going to school ... ... ... 
\\ \hline
Gold & 
i lied to a new coworker about dead sister being alive, dads the boss.
\\ \hline
Model & 
don't lie to your boss.
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Reddit TIFU sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{Reddit TIFU} \\ \hline
\makecell{Document\ID \#224)} & 
so this happened last night and proceeded into today. went out with this girl for drinks, we started fooling around. went back to my car and she went ape-shit on my neck. didn't notice anything at the time, woke up in the morning with a nice bruise on my neck. went to work in a very corporate setting. thought that my dress shirt buttoned all the way up would be enough to conceal it but nope. boss inquires about the bruise, i say "yeah i don't know where i got that, they keep popping up all over my body". i then proceeded to show him bruises on my arms (i was trying to climb trees whilst drunk over the weekend) as a cover up. boss became concerned and said i need to see a doctor right away as it could be something more serious like leukemia. he has given me next monday off to see the doctor...
\\ \hline
Gold & 
girl attacked my neck, boss sent me to the doctor.
\\ \hline
Model & 
went out with a girl, she ape-shit on my neck, woke up with a bruise, tried to cover it up by showing boss bruises on my arms.
\\ \hline
ROUGEL-F1 & 
21.05
\\ \hline
\makecell{Document\ID \#175)} & 
regret and guilt are two of the worst things that one can feel. unlike most tifu's this one actually occurred today. me and my little brother who is 10 years old were playing football. sometimes my little brother is really annoying and this was a prime case. he decided to kick my shin, now i'm notorious for being quick to anger and can be violent, i know not one of my better traits. so i pushed him and to be honest i pushed him pretty hard. he fell and presumably stuck out his left arm to break his fall. however he instead landed on his wrist. at first it looked like he was ok but then he cried out in pain and screamed and shouted very loudly. at first i though he was bluffing because he often was melodramatic and did things like this. however after 5 minutes it became apparent that he was completely serious. i lifted him and sat him down on a soft chair. since i have done first aid training i applied ice to his wrist. then i called my mum and told her everything including that it was all my fault. i decided that lying wasn't my best option and just spilt the beans and told her everything. she was really pissed off and told us that she would be home in ten minutes. when i heard the knock on the door i knew it was not going to be good. she was really angry and shouted at me before looking at my brother. she saw his wrist and turned pale. it was black or horrible looking like something you could find on google images but it was definitely swollen. she decided to take him to a\&e, on the way out she grilled me about ... ... ... 
\\ \hline
Gold & 
i managed to accidentally break my brothers arm. my aims to buy a car before uni have faced a major setback and my hopeful future in a medical career has faced a major setback and it's all my fault.
\\ \hline
Model & 
i pushed my brother and he broke his wrist.
\\ \hline
ROUGEL-F1 & 
12.24
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on Reddit TIFU sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{BIGPATENT} \\ \hline
\makecell{Document\ID \#298)} & 
it will be readily apparent to one skilled in the art that various substitutions and modifications may be made in the invention disclosed herein without departing from the scope and spirit of the invention . the term “ a ” or “ an ” as used herein in the specification may mean one or more . as used herein in the claim ( s ), when used in conjunction with the word “ comprising ”, the words “ a ” or “ an ” may mean one or more than one . as used herein “ another ” may mean at least a second or more . the term “ animal ” as used herein refers to any species of the animal kingdom . in preferred embodiments it refers more specifically to humans , animals in their wild state , animals used as pets ( birds , dogs , cats , horses ), animals used for work ( horses , cows , dogs ) and animals which produce food ( chickens , cows , fish ), farm animals ( pigs , horses , cows , sheep , chickens ) or are themselves food ( frogs , chickens , fish , crabs , lobsters , shrimp , mussels , scallops , goats , boars , cows , lambs , pigs , ostrich , emu , eel ) and other animals well known to the art . the term “ effective amount ” as used herein is defined as the amount of the composition required to produce an effect in a host which can be monitored using several endpoints known to those skilled in the art . in a specific embodiment , these endpoints are surrogate markers . the term “ feed conversion efficiency ” as used herein is defined as the ... ... ... 
\\ \hline
Gold & 
the intramuscular electroporated injection of a protease - resistant growth hormone - releasing hormone cdna into rat dams at 16 days of gestation resulted in the enhanced long - term growth of the f1 offspring . the offspring were significantly heavier by one week of age and the difference was sustained to 10 weeks of age . consistent with their augmented growth , plasma igf - i concentration of the f1 progeny was increased significantly . the pituitary gland of the offspring was significantly heavier , and contained an increased number of somatotropes and lactotrophs , and is indicative of an alteration in cell lineages . these unique findings demonstrate that enhanced ghrh expression in pregnant dams can result in intergenerational growth promotion , by altering development of the pituitary gland in the offspring .
\\ \hline
Model & 
the present invention relates to the use of growth hormone releasing hormone and growth hormone secretagogue receptors in the manufacture of medicaments for the treatment of growth deficiencies in newborn animals .
\\ \hline
ROUGE1-F1 & 
22.22
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on BIGPATENT sampled by ROUGE1-F1.}
\end{table*} \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{BIGPATENT} \\ \hline
\makecell{Document\ID \#155)} & 
the automated edge finishing apparatus and method of the present invention comprises hardware components and associated software , providing the ability to learn a trajectory of a workpiece by \&\# 34 ; guarded moves ,\&\# 34 ; machining the workpiece by moving the workpiece along the trajectory and controlling the trajectory in real time . similar workpieces can be machined from the learned trajectory without expensive fixtures by the sensors and software locating the tool and adjusting the trajectory . the invention is useful for performing edge finishing on intricately designed workpieces or workpieces with close edge finishing tolerances . the configuration of the preferred system is shown in fig1 and 2 . the preferred invention comprises an xy table 10 , a force sensor 12 , cutting tool 14 , xy table position control hardware 32 , and a control system as shown in fig2 . the preferred xy table 10 comprises a daedal series 200000 , or the like , powered by tlm120 linear motors from trilogy systems corporation , or the like . the linear actuators comprise three - phase brushless linear motors , with a stationary magnet track and a moving coil assembly . the linear motors are driven by dc amplifiers , such as pwm amplifiers . both table axes are equipped with incremental encoders with a resolution of 12 , 700 counts / inch . the innermost level of table control is position control , by a controller 32 such as a dmc - 520 motion controller from galil motion control , as shown in fig2 . this board controls position of both table axes using proportional plus derivative ( pd ) control action . the resulting position - controlled table is the starting point for the force control system discussed below . a ... ... ... 
\\ \hline
Gold & 
the disclosure is directed to an apparatus and method for automated edge finishing using hybrid position / force control of an xy table . the disclosure is particularly directed to learning the trajectory of the edge of a workpiece by \&\# 34 ; guarded moves \&\# 34 ;. machining is done by controllably moving the xy table , with the workpiece mounted thereon , along the learned trajectory with feedback from a force sensor . other similar workpieces can be mounted , without a fixture on the xy table , located and the learned trajectory adjusted
\\ \hline
Model & 
an apparatus and method for edge finishing comprising a position - controlled linear table , a cutting tool , a force sensor , and a control system . the position - controlled table allows the workpiece to be moved past the cutting tool . the table is equipped with two linear motors and associated encoders for position control . the control system can be programmed to allow the workpiece to be cut from a learned trajectory in real time . the cutting tool can be programmed to cut the workpiece along the trajectory , or it can be programmed to cut the workpiece on a normal path . the system combines the two major elements of edge finishing : position control and force control . the position control allows the workpiece to be moved past the cutting tool , while the force control allows the workpiece to be moved in the normal direction . the system components can be placed in modular form , allowing rapid upgrade and replacement of existing systems .
\\ \hline
ROUGE2-F1 & 
11.72
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on BIGPATENT sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{BIGPATENT} \\ \hline
\makecell{Document\ID \#237)} & 
referring now to fig2 - 4 a preferred embodiment of the control system 10 of the present invention is shown mounted upon the underside of gun 11 and comprised of an emd 12 , electronic circuitry denoted schematically by box 13 , a battery 14 , and electrical switch trigger 15 . the illustrated gun is comprised of a forestock 22 which supports barrel 16 , receiver portion 17 located at the rear extremity of said barrel , magazine 18 , conventional trigger 19 with associated pistol grip 20 , and shoulder stock 21 which contains a compressed propellant gas . the emd is intended to produce reciprocating linear movement of a push rod 23 . the emd may be a solenoid , either of an in - line type or clapper or rotary type . alternatively , the emd may be a servo type device using an arm , lever or gear system to activate rod 23 . all such devices are characterized in that a pulse of electrical energy produces a controlled mechanical force , and the discontinuation of said pulse either produces a reverse force or permits interaction of a reverse force such as may be produced by a spring - biased conventional trigger . the emd may be secured to forestock 22 by brackets or removable fastening means . a push rod 23 , extending from said emd is slideably positioned by guide 25 mounted by bracket 35 beneath receiver portion 17 . in those embodiments wherein the emd is a solenoid , guide 25 may not be required . the length of rod 23 and its positioning by guide 25 is such as to cause the distal extremity 26 of rod 23 to contact trigger 19 . said distal extremity may be equipped with coupling ... ... ... 
\\ \hline
Gold & 
a system for controlling the firing of a paint ball gun includes an electromechanical device that produces linear motion interactive with the conventional trigger of the gun . electronic circuitry produces pulses that activate the emd . the pulses are based upon control signals produced by either the timing of paint balls entering the firing chamber of the gun or the detection of the presence of a single paint ball within the chamber . a manually operated electrical switch trigger activates the electronic circuitry . an adjustably predetermined number of paint balls will fire based upon each depression of the electrical switch trigger while minimizing the chopping of paint balls in the firing chamber .
\\ \hline
Model & 
a paint ball gun control system which permits selective firing of paint balls in response to the depression of the trigger . the system includes an electronic circuit which converts a dc current into a regulated pulse current , and an electromagnetic device which is actuated by the regulated pulse current to produce a reciprocating linear motion . a push rod attached to the emd is adapted to contact the trigger of the gun .
\\ \hline
ROUGEL-F1 & 
22.10
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on BIGPATENT sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{arXiv} \\ \hline
\makecell{Document\ID \#289)} & 
machine learning methods are used widely within high energy physics ( hep ) . one promising approach , used extensively outside of hep for applications such as handwriting recognition , is that of support vector machines ( svms ) , a supervised learning model used with associated learning algorithms for multivariate analysis ( mva ) . developed originally in the 1960s , with the current standard version proposed in 1995 \@xcite , svms aim to classify data points using a maximal margin hyperplane mapped from a linear classification problem to a possibly infinite dimensional hyperspace . however this means svms , like other mva classifiers , have a number of free parameters which need to be tuned on a case by case basis . this motivates a number methods for ensuring the classifier is sufficiently generalised such that when used on an unseen dataset the performance can be accurately predicted . in this paper a brief overview of svms is given in section [ sec : svm ] , with an example using svms shown in section [ sec : checker ] . generalisation is discussed in section [ sec : gen ] with an illustrative example of how this can improve performance given in section [ sec : check2 ] . consider the problem of linear classification with the svm where the training set , \@xmath0 , is linearly separable . we define a separating hyperplane given by \@xmath1 , where \@xmath2 , the weight vector , is perpendicular to the hyperplane , and \@xmath3 , the bias , determines the distance of the hyperplane from the origin ( fig . ) . a decision function defined by \@xmath4 is used to positively and negatively classify \@xmath5 , the points in the training set . without further constraint the ... ... ... 
\\ \hline
Gold & 
we review the concept of support vector machines ( svms ) and discuss examples of their use . one of the benefits of svm algorithms , compared with neural networks and decision trees is that they can be less susceptible to over fitting than those other algorithms are to over training . this issue is related to the generalisation of a multivariate algorithm ( mva ) ; a problem that has often been overlooked in particle physics . we discuss cross validation and how this can be used to improve the generalisation of a mva in the context of high energy physics analyses . the examples presented use the toolkit for multivariate analysis ( tmva ) based on root and describe our improvements to the svm functionality and new tools introduced for cross validation within this framework .
\\ \hline
Model & 
support vector machines ( svms ) are one of the most promising methods for machine learning in high energy physics . svms aim to classify data points using a maximal margin hyperplane mapped from a linear classification problem to a possibly infinite dimensional hyperspace . however this means svms , like other mva classifiers , have a number of free parameters which need to be tuned on a case by case basis . this motivates a number methods for ensuring the classifier is sufficiently generalised such that when used on an unseen dataset the performance can be accurately predicted . in this paper a brief overview of svms is given , with an example using svms shown in section [ sec : checker ] . generalisation is discussed with an illustrative example of how this can improve performance given in section [ sec : gen ] .
\\ \hline
ROUGE1-F1 & 
35.29
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on arXiv sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{arXiv} \\ \hline
\makecell{Document\ID \#32)} & 
deep neural networks ( dnns ) have been receiving ubiquitous success in wide applications , ranging from computer vision \@xcite , to speech recognition \@xcite , natural language processing \@xcite , and domain adaptation \@xcite . as the sizes of data mount up , people usually have to increase the number of parameters in dnns so as to absorb the vast volume of supervision . high performance computing techniques are investigated to speed up dnn training , concerning optimization algorithms , parallel synchronisations on clusters w / o gpus , and stochastic binarization / ternarization , etc \@xcite . on the other hand the memory and energy consumption is usually , if not always , constrained in industrial applications \@xcite . for instance , for commercial search engines ( e.g. , google and baidu ) and recommendation systems ( e.g. , netflix and youtube ) , the ratio between the increased model size and the improved performance should be considered given limited online resources . compressing the model size becomes more important for applications on mobile and embedded devices \@xcite . having dnns running on mobile apps owns many great features such as better privacy , less network bandwidth and real time processing . however , the energy consumption of battery - constrained mobile devices is usually dominated by memory access , which would be greatly saved if a dnn model can fit in on - chip storage rather than dram storage ( c.f . \@xcite for details ) . a recent trend of studies are thus motivated to focus on compressing the size of dnns while mostly keeping their predictive performance \@xcite . with different intuitions , there are mainly two types of dnn compression methods , which could be used in conjunction for better parameter savings . the ... ... ... 
\\ \hline
Gold & 
as the complexity of deep neural networks ( dnns ) trend to grow to absorb the increasing sizes of data , memory and energy consumption has been receiving more and more attentions for industrial applications , especially on mobile devices . this paper presents a novel structure based on functional hashing to compress dnns , namely funhashnn . for each entry in a deep net , funhashnn uses multiple low - cost hash functions to fetch values in the compression space , and then employs a small reconstruction network to recover that entry . the reconstruction network is plugged into the whole network and trained jointly . funhashnn includes the recently proposed hashednets \@xcite as a degenerated case , and benefits from larger value capacity and less reconstruction loss . we further discuss extensions with dual space hashing and multi - hops . on several benchmark datasets , funhashnn demonstrates high compression ratios with little loss on prediction accuracy .
\\ \hline
Model & 
deep neural networks ( dnns ) have been receiving ubiquitous success in wide applications , ranging from computer vision to speech recognition . as the sizes of data mount up , people usually have to increase the number of parameters in dnns so as to absorb the vast volume of supervision . compressing the model size becomes more important for applications on mobile and embedded devices . with different intuitions , there are mainly two types of dnn compression methods , which could be used in conjunction for better parameter savings . the first type tries to revise the training target into more informative supervision using dark knowledge . the second type observes the redundancy existence in network weights , and exploits techniques to constrain or reduce the number of free - parameters in dnns during learning . in applications , we observe hashednets compresses model sizes greatly at marginal loss of accuracy for some situations , whereas also significantly loses accuracy for others . after revisiting its mechanism , we conjecture this instability comes from at least three factors . first , hashing and training are disjoint in a two - phase manner , i.e. , once inappropriate collisions exist , there may be no much optimization room left for training . second , one single hash
\\ \hline
ROUGE2-F1 & 
7.21
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on arXiv sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{arXiv} \\ \hline
\makecell{Document\ID \#270)} & 
the alice experiment is dedicated to the study of the properties of qcd matter created in nucleus - nucleus collisions at lhc energies \@xcite . the inner tracking system in the alice apparatus is made of position sensitive detectors which have to handle several thousands tracks per unit of rapidity . the two innermost layers at 3.9 \@xmath10 and 7.6 \@xmath10 radii , respectively , constitute the silicon pixel detector ( spd ) . the spatial precision and hit efficiency of the spd are key parameters since they determine the alice capability of detecting particles with open heavy - flavour \@xcite . + the basic detector unit of the alice spd is the ladder , a two - dimensional silicon matrix of p\@xmath11n reverse biased diodes of dimensions 50 x 425 \@xmath12 , flip - chip bonded to five read - out chips . each diode is connected to a cell of the front - end read - out asic via a pb - sn solder bump of 25 \@xmath13 diameter . the detector contains nearly 10\@xmath14 active cells in total . the read - out is binary . to reduce the material budget , the sensor thickness is limited to 200 \@xmath13 and the read - out chip wafers are thinned down to 150 \@xmath13 . further details can be found in \@xcite . + early prototypes of the alice spd elements , in the form of single - chip assemblies , were tested in high energy proton / pion beams at the cern sps in 2002 and 2003 . these assemblies were made with sensors of 200 \@xmath13 and 300 \@xmath13 thicknesses , while the read - out chips ( unthinned ) were 725 \@xmath13 thick . those beam tests were primarily aimed at evaluating the performance of ... ... ... 
\\ \hline
Gold & 
the two innermost layers of the alice inner tracking system are instrumented with silicon pixel detectors . single chip assembly prototypes of the alice pixels have been tested in high energy particle beams at the cern sps . detection efficiency and spatial precision have been studied as a function of the threshold and the track incidence angle . the experimental method , data analysis and main results are presented . d. elia\@xmath0 , g. anelli\@xmath1 , f. antinori\@xmath2 , a. badal\@xmath3 , g.e . bruno\@xmath4 , m. burns\@xmath1 , i.a . cali\@xmath5 , m. campbell\@xmath1 , m. caselle\@xmath4 , s. ceresa\@xmath1 , p. chochula\@xmath1 , m. cinausero\@xmath6 , j. conrad\@xmath1 , r. dima\@xmath2 , d. fabris\@xmath2 , r.a . fini\@xmath4 , e. fioretto\@x
\\ \hline
Model & 
the spatial precision and hit efficiency of the alice silicon pixel detector ( spd ) are key parameters for the detection of heavy - flavour in nucleus - nucleus collisions at lhc energies . the performance of the detector under test , in the form of single - chip assemblies , together with a detailed cluster analysis of the hits , can yield a good determination of the intrinsic spatial precision and detection efficiency of the pixel plane under test . results from the 2002 beam test at the cern sps , where the sensor thickness was the same as the one used in ladder production , are presented and compared with the main results for the thicker sensor . , , , , , , , ,
\\ \hline
ROUGEL-F1 & 
17.70
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on arXiv sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{PubMed} \\ \hline
\makecell{Document\ID \#124)} & 
acute appendicitis is considered as the most frequent cause of acute abdomen in young patients with a slight predominance in males . although its clinical diagnosis is simple , 20 - 30\% of appendicitis cases are reported to be gangrenous or perforated because of delayed diagnosis , which can lead to increased morbidity and mortality . additionally , up to 10\% of patients especially older ones , immunocompromised patients , and pregnant women are diagnosed with atypical presentations such as appendiceal mass which requires long hospital admissions , intravenous antibiotics , and often needs percutaneous drainage . using colonoscopy for the diagnosis and treatment of asymptomatic acute appendicitis and appendiceal abscess are rare . in unusual appendicitis cases , especially in the presence of abdominal comorbidities , colonoscopy is used ; although colonoscopy is not considered as the first line intervention because of complications risk such as bleeding or perforation . furthermore non - invasive tests such as ultrasonography and computed tomography with more sensitivity can be used . we report an atypical case of appendiceal abscess , which was diagnosed during colonoscopy . a 30-year - old man referred for evaluation to our center with epigastric pain from three days prior to his evaluation . the pain was nonspecific and colicky distributed in periumblical region , did not radiate to any site and was unrelated to bowel movements . he had experienced also nausea , anorexia and intermittent vomiting . there was a history of marijuana and alcohol usage from several years ago . upon physical examination , he was ill with restlessness . other laboratory values such as hemoglobin , platelets , serum amylase , erythrocyte sedimentation rate , aspartate aminotransferase , alanine aminotransferase , alkaline phosphatase , and electrolytes were normal . his workup on admission including ... ... ... 
\\ \hline
Gold & 
although complicated acute appendicitis is occasionally observed in general practice , the diagnosis of appendiceal abscess by colonoscopy is unusual . we report a case of appendiceal abscess with spontaneous drainage into the lumen of colon diagnosed during colonoscopy and resolved by the time of surgery .
\\ \hline
Model & 
acute appendicitis is considered as the most frequent cause of acute abdomen in young patients with a slight predominance in males . although its clinical diagnosis is simple , 20 - 30\% of appendicitis cases are reported to be gangrenous or perforated because of delayed diagnosis , which can lead to increased morbidity and mortality . additionally , up to 10\% of patients especially older ones , immunocompromised patients , and pregnant women are diagnosed with atypical presentations such as appendiceal mass which requires long hospital admissions , intravenous antibiotics , and often needs percutaneous drainage . using colonoscopy for the diagnosis and treatment of asymptomatic acute appendicitis and appendiceal abscess are rare . in unusual appendicitis cases , especially in the presence of abdominal comorbidities , colonoscopy is used ; although colonoscopy is not considered as the first line intervention because of complications risk such as bleeding or perforation . we report an atypical case of appendiceal abscess , which was diagnosed during colonoscopy .
\\ \hline
ROUGE1-F1 & 
31.25
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on PubMed sampled by ROUGE1-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{PubMed} \\ \hline
\makecell{Document\ID \#72)} & 
this retrospective study was approved by the institutional review board of our hospital ( 2014 - 09 - 037 ) . between november 2011 and october 2014 , 245 patients in the icu underwent bedside pdt with a ciaglia blue rhino percutaneous tracheostomy set ( cook medical , bloomington , in , usa ) due to prolonged endotracheal tube insertion . pdt was not performed in patients aged under 18 years , those with any pulsation palpated over the tracheostomy site , patients with a history of surgery or radiotherapy in the cervical region , and those with coagulopathy ( increased prothrombin time , inr > 2 ) . the first four procedures were performed by two physicians and one nurse using the standard technique , with one physician operating the bronchoscope while the other performed pdt . the next 55 procedures were performed by one physician and one nurse , with the physician using the bronchoscope only to confirm and evaluate the proper depth for the endotracheal tube before performing pdt using the simplified method described below . the remaining 186 procedures were performed using the simplified technique after repositioning the endotracheal tube at the predetermined depth ( 17 cm for males and 15 cm for females at the incisor ) . all pdts were performed by a single physician and one nurse , both of whom fully understood the entire procedure . the nurse administered sedatives and analgesics , as well as repositioning the endotracheal tube . except for patients with limited neck extension , a small pillow was placed under the patient 's shoulders to slightly extend the neck . transverse markings were made at the cricothyroid membrane and 2.0 cm toward the sternal notch ( fig . was also made in the midline of the trachea ... ... ... 
\\ \hline
Gold & 
when first introduced , percutaneous dilatational tracheostomy ( pdt ) was performed using a bronchoscope . the bronchoscope itself , however , has several disadvantages , including interruption of ventilation . modifications of the standard pdt technique have suggested that pdt can be safely performed without a bronchoscope . however , the safety of these modifications is unclear and many modifications have yet to be widely adopted . this study retrospectively evaluated the safety of a simplified pdt technique using the cook ciaglia blue rhino in 186 patients . after confirming the underlying structures by palpation , pdt was performed without a bronchoscope or blunt dissection , 2.0 cm below the cricothyroid membrane . ultrasonography was used only when palpation was difficult . except for bleeding in three patients with coagulopathy , none of the patients experienced any early or late complications , and there were no deaths related to pdt . these findings confirm that pdt can be safely performed using a simplified technique , which may be useful in situations with limited equipment and medical personnel .
\\ \hline
Model & 
backgroundpercutaneous tracheostomy ( pdt ) is one of the most common tracheostomy procedures performed in the intensive care unit ( icu ) . although the use of a fiberoptic bronchoscope ( fob ) during pdt has been recommended for precise positioning and to avoid complications , bronchoscopy itself can cause complications . this study assessed the safety of a simplified pdt technique using the ciaglia blue rhino , by repositioning the existing endotracheal tube at a pre - determined depth and determining proper pdt site by superficial palpation.methodsbetween november 2011 and october 2014 , 245 patients in the icu underwent bedside pdt with a ciaglia blue rhino percutaneous tracheostomy set due to prolonged endotracheal tube insertion . the first four procedures were performed by two physicians and one nurse using the standard technique , with one physician operating the bronchoscope while the other performed pdt . the next 55 procedures were performed by one physician and one nurse , with the physician using the bronchoscope only to confirm and evaluate the proper depth for the endotracheal tube before performing pdt using the simplified method described below . the remaining 186 procedures were performed using the simplified technique after repositioning the endo
\\ \hline
ROUGE2-F1 & 
10.53
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on PubMed sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{PubMed} \\ \hline
\makecell{Document\ID \#114)} & 
chronic pain affects up to 20\% of the population in developed nations.14 this represents a profound impact on individuals and their families alongside the sizeable burden on employers , health care systems , and society in general.3 when chronic pain occurs , it has the potential to become disease itself , and subsequently , chronic pain has emerged as a distinct phenomenon.5 management of chronic pain varies greatly between nations and even within nations . literature supports a multidisciplinary approach as the standard of care , although various health care systems may not always support this concept consistently.2 the current standard of care for chronic , noncancer pain typically includes many disciplines with the clinician developing an individualized treatment plan with the options of utilizing surgical interventions , pharmacology , and psychological and physical therapies . opioid analgesics are often prescribed , despite the lack of clinical evidence supporting their long - term use in the management of chronic pain.6 however , for many patients , this multidisciplinary approach is inadequate or ineffectual or is accompanied by the burden of side effects that are unacceptable and debilitating . only at this late stage , the field of neuromodulation for the treatment of pain has developed rapidly since the seminal paper on the electrical inhibition of pain by the stimulation of the dorsal column almost 50 years ago.7 the original term of dorsal column stimulation has evolved to become known as spinal cord stimulation ( scs).8 scs has been particularly effective as an adjunct in treating mixed neuropathic / nociceptive and neuropathic / radicular pain conditions such as failed back surgery syndrome ( fbss ) and complex regional pain syndrome ( crps ) . neuromodulation therapies offer a treatment option that has minimal side effects and that is relatively safe and ... ... ... 
\\ \hline
Gold & 
spinal cord stimulation ( scs ) applications and technologies are fast advancing . new scs technologies are being used increasingly in the clinical environment , but often there is a lag period between the clinical application and the publishing of high - quality evidence on safety and efficacy . recent developments will undoubtedly expand the applicability of scs , allowing more effective and individualized treatment for patients , and may have the potential to salvage patients who have previously failed neuromodulation . already , high - level evidence exists for the safety , efficacy , and cost - effectiveness ( level i ii ) of traditional scs therapies in the treatment of chronic refractory low back with predominant limb pain ( regardless of surgical history ) . more than half of all patients with chronic painful conditions experience sustained and significant levels of pain reduction following scs treatment . although only limited evidence exists for burst stimulation , there is now level i evidence for both dorsal root ganglion scs and high - frequency scs that demonstrates compelling results compared with traditional therapies . the body of evidence built on traditional scs research may be redundant , with newer iterations of scs therapies such as dorsal root ganglion scs , high - frequency scs , and burst scs . a number of variables
\\ \hline
Model & 
chronic pain affects up to 20\% of the population in developed nations . the current standard of care for chronic , noncancer pain typically includes many disciplines with the clinician developing an individualized treatment plan with the options of utilizing surgical interventions , pharmacology , and psychological and physical therapies . however , for many patients , this multidisciplinary approach is inadequate or ineffectual or is accompanied by the burden of side effects that are unacceptable and debilitating . only at this late stage , the field of neuromodulation for the treatment of pain has developed rapidly since the seminal paper on the electrical inhibition of pain by the stimulation of the dorsal column almost 50 years ago . spinal cord stimulation ( scs ) has been particularly effective as an adjunct in treating mixed neuropathic / nociceptive and neuropathic / radicular pain conditions such as failed back surgery syndrome ( fbss ) and complex regional pain syndrome ( crps ) . in traditional scs therapies , the objective has been to replace the pain sensation with paresthesia that requires mapping of stimulation to the region of pain . in recent years , the next phase in the evolution of neuromodulation has become available with the development of dorsal root ganglion scs and the emerging use of two novel advances in stimulation frequencies
\\ \hline
ROUGEL-F1 & 
17.68
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on PubMed sampled by ROUGEL-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{AESLC} \\ \hline
\makecell{Document\ID \#33)} & 
The funerals of Andy, Darren and Paul, who tragically lost their lives in last Wednesday's explosion at Teesside Power Station, will be held this week at the following times: Andy Sherwood 2.00pm, Wednesday, August 15 St James Church, Hartlepool Darren Higgins 9.15 am, Thursday, August 16 St. Andrews Church, Teesville, Middlesbrough Paul Surtees 1.30 pm, Thursday, August 16 Easington Church of the Ascension, Easington Colliery We know that everyone at Enron will want to remember the colleagues we have lost at this difficult time, and our thoughts and prayers are with their families and friends. We are considering how best to commemorate the lives of Andy, Darren and Paul, and will be discussing our ideas with their families, in due course. Graeme White is recovering in Middlesbrough General Hospital where his condition remains stable and we are continuing to provide as much support as possible for Graeme and his family. The plant remains off-line and Enron continues to co-operate fully with the Health \& Safety Executive's investigation into the cause of the explosion. John Sherriff and Michael Brown
\\ \hline
Gold & 
TEESSIDE POWER STATION -- UPDATE
\\ \hline
Model & 
Teesside Update
\\ \hline
ROUGE1-F1 & 
66.67
\\ \hline
\makecell{Document\ID \#231)} & 
Steve: I just spoke with Bill Bradford and I understand that Bill and Mark Tarkington discussed the use of "Specified Entity" and agreed, subject to DB committee approval, to eliminate the use of Specified Entity in the ENA and affiliated masters with DB (i.e., "Specified Entity: none"). Once you have confirmed this with Mark, please prepare a final blacklined draft for our review and then execution copies of the ISDA Master Agreement. In connection with the final draft, Tana Jones will fax to you minor format comments. [She will also fax to you the electronic trading agreements and password applications for London and U.S. based trading.] The following ENA affiliates will also need to execute ISDA Master Agreements with DB: (1) ECT Investments, Inc. (Delaware corporation) (2) Enron Capital \& Trade Resources International Corp. (Delaware corporation) (3) EnronCredit.com Limited (U.K. corporation)
\\ \hline
Gold & 
Enron North America Corp. ("ENA")/Deutsche Bank ("DB") ISDA Master Agreement
\\ \hline
Model & 
Deutsche Bank
\\ \hline
ROUGE1-F1 & 
30.77
\\ \hline
\makecell{Document\ID \#108)} & 
Due to a variety of circumstances we did not migrate employees to the UBSWE environment tonight (Monday, February 4, 2002). We will communicate further details of the migration plan on Tuesday, February 5, 2002. Thank you for your cooperation in advance, please contact the Resolution Center at x 3-1411 should you have any questions. UBSW Energy Global Infrastructure
\\ \hline
Gold & 
DRAFT: INFORMATION, Please Read: UBSWE Migration postponed
\\ \hline
Model & 
migrated employees
\\ \hline
ROUGE1-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on AESLC sampled by ROUGE1-F1.}
\end{table*} \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{AESLC} \\ \hline
\makecell{Document\ID \#100)} & 
El Paso recently filed their proposal to reallocate receipt point capacity. Comments are due on Thursday, May 17, 2001. We would like to meet with you to make sure our comments address the commercial concerns regarding the reallocation plan. In that regard, please plan to attend a meeting on Thursday, May 3 at 3:45 in 3269. Becky will provide for your review a summary of El Paso's proposal. Please let me know if you have any questions. Thanks
\\ \hline
Gold & 
El Paso's Proposal to Address the Allocation of Receipt Point Capacity
\\ \hline
Model & 
Comments on El Paso's Reallocation Proposal
\\ \hline
ROUGE2-F1 & 
23.53
\\ \hline
\makecell{Document\ID \#189)} & 
Nothing from Ben yet. If I don't hear from him in the next couple of hours, then I think we should be ready with an extension letter, giving them a couple of days extra for free. Ben indicated to me last week that he would give them some time once they asked. MK called with questions on the assignment - nothing serious. Kay
\\ \hline
Gold & 
NW
\\ \hline
Model & 
SMUD
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\makecell{Document\ID \#37)} & 
Kal/Susan - Attached is an updated version of the disclosure consent agreement for EOL information. Please take a look at the attached and let me know your thoughts. Mark
\\ \hline
Gold & 
EOL Disclosure Consent
\\ \hline
Model & 
EOL Consent
\\ \hline
ROUGE2-F1 & 
0.00
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on AESLC sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{AESLC} \\ \hline
\makecell{Document\ID \#106)} & 
As discussed during our recent demonstration of our new portfolio tracking system (PortRAC), we have completed a security system that will provide a "firewall" controlling access to each business unit's data. Our intent is to open up the system to designated members of each business unit or others as approved by the COO or Managing Director of that group. In order to complete the initial setup, we need to receive a listing of approved users for your business unit and their login Id's. If you prefer to delegate security approval authority to someone else on your staff, please let us know and we will establish procedures with them. If additional demonstrations of PortRAC are necessary for your group, please let us know. Thank you for your assistance. Rick C.
\\ \hline
Gold & 
PortRAC Security
\\ \hline
Model & 
PortRAC Security System
\\ \hline
ROUGEL-F1 & 
80.00
\\ \hline
\makecell{Document\ID \#267)} & 
Susan: I'm not sure that what I told you to do in the CSA Annex with respect to Custodians and their qualifications work so for now let's keep it out and go back to our old language. Sara and Mark: Shari had a counterparty point out to her that in our CSA Annex language where we talk about the circumstances under which a party can "hold" collateral, all of our triggers are based on something happening to our counterparty and nothing is said about what would happen if a Custodian holds the colateral and fails to be creditworthy. Do you have any thoughts on this? At first I thought that we should require the Custodian to be a Qualified institution but I'm not sure that that is the right approach. Let me know what you think. Carol
\\ \hline
Gold & 
Custodian Language
\\ \hline
Model & 
Custodians
\\ \hline
ROUGEL-F1 & 
0.00
\\ \hline
\makecell{Document\ID \#22)} & 
the tax issue that we discussed at the rmt meeting has been cleared. now we need to see if the \\text{PEGASUS}_{\text{LARGE}}8,000 or the portion of the purchase price of the principal residence attributable such fixtures or movable facilities.''. (c) Eligible Veteran.-- (1) In general.--Section 36(c)(1) of such Code is amended by striking   First-time homebuyer.--The term   first time homebuyer' means any individual'' and inserting   Eligible veteran.--The term   eligible veteran' means any individual who is a veteran (as defined in section 101(2) ... ... ... 
\\ \hline
Gold & 
Veterans Homebuyer Accessibility Act of 2015 This bill amends the Internal Revenue Code, with respect to the tax credit for first-time homebuyers, to: (1)\&nbsp;allow a veteran of the Armed Forces a tax credit for 10\% of the purchase price of a principal residence purchased prior to January 1, 2018; (2)\&nbsp;allow an additional credit for the cost, not exceeding \\text{PEGASUS}_{\text{LARGE}}150 million from the Supplemental Medical Insurance Trust Fund to the CMS Program Management Account.
\\ \hline
Model & 
Medicare Common Access Card Act of 2017 This bill requires the Centers for Medicare \&amp; Medicaid Services (CMS) to establish a pilot program that demonstrates the feasibility of using smart card technology to address Medicare fraud. Under the program, smart card technology must be issued free-of-charge to selected Medicare beneficiaries, suppliers, and providers; such technology must support the secure, electronic authentication of beneficiary identity at points of service. In selecting program participants, the CMS must consider the risk of fraud, waste, or abuse among categories of suppliers and providers.
\\ \hline
ROUGE2-F1 & 
25.10
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on BillSum sampled by ROUGE2-F1.}
\end{table*}
 \begin{table*}[bth!]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c|p{0.85\textwidth}}
\hline
\multicolumn{2}{c}{BillSum} \\ \hline
\makecell{Document\ID \#216)} & 
SECTION 1. SHORT TITLE; TABLE OF CONTENTS. (a) Short Title.--This Act may be cited as the   Meeting Our Responsibility to Medicare Beneficiaries Act of 2005''. (b) Table of Contents.--The table of contents of this Act is as follows: Sec. 1. Short title; table of contents. TITLE I--ELIMINATING SPECIAL INTEREST PREFERENCES Sec. 101. Negotiating fair prices for medicare prescription drugs. Sec. 102. Elimination of MA Regional Plan Stabilization Fund (Slush Fund). Sec. 103. Application of risk adjustment reflecting characteristics for the entire medicare population in payments to Medicare Advantage organizations. TITLE II--IMPROVING THE MEDICARE PROGRAM FOR BENEFICIARIES Sec. 201. Eliminating coverage gap. Sec. 202. Requiring two prescription drug plans to avoid Federal fallback. Sec. 203. Waiver of part D late enrollment penalty for transition period. Sec. 204. Improving the transition of full-benefit dual eligible individuals to coverage under the medicare drug benefit. Sec. 205. Part B premium reduction. Sec. 206. Study and report on providing incentives to preserve retiree coverage. Sec. 207. Promoting transparency in employer subsidy payments. TITLE I--ELIMINATING SPECIAL INTEREST PREFERENCES SEC. 101. NEGOTIATING FAIR PRICES FOR MEDICARE PRESCRIPTION DRUGS. (a) In General.--Section 1860D-11 of the Social Security Act (42 U.S.C. 1395w-111) is amended by striking subsection (i) (relating to noninterference) and by inserting the following new subsection:   (i) Authority To Negotiate Prices With Manufacturers.--   (1) In general.--The Secretary shall have authority similar to that of other Federal entities that purchase prescription drugs in bulk to negotiate contracts with manufacturers of covered part D drugs, consistent with the requirements and in furtherance of the goals of providing quality care and containing costs under this part.   (2) Required use of authority.--   (A) Fallback plans.--The Secretary shall exercise the authority described in paragraph (1) with respect to covered part D drugs ... ... ... 
\\ \hline
Gold & 
Meeting Our Responsibility to Medicare Beneficiaries Act of 2005 - Amends title XVIII (Medicare) of the Social Security Act (SSA) with respect to: (1) negotiating fair prices for Medicare prescription drugs; (2) elimination of the MA Regional Plan Stabilization Fund; (3) application of risk adjustment reflecting characteristics for the entire Medicare population in payments to Medicare advantage organizations; (4) modification of the annual out-of-pocket threshold with respect to prescription drug benefits; (5) requiring two prescription drug plans to avoid Federal fallback; (6) waiver of the part D (Voluntary Prescription Drug Benefit Program); (7) transition of full-benefit dual eligible individuals to coverage under the Medicare drug benefit; and (8) Medicare part B (Supplementary Medical Insurance) premium reduction. Directs the Secretary to study and report to Congress on providing incentives to preserve retiree coverage. Amends SSA title XVIII to direct the Secretary to make certain information regarding the sponsor of a qualified prescription drug plan receiving a subsidy under the prescription drug program available to the public through the Internet website of the Centers for Medicare \& Medicaid Services.
\\ \hline
Model & 
Meeting Our Responsibility to Medicare Beneficiaries Act of 2005 - Amends title XVIII (Medicare) of the Social Security Act to repeal the prohibition against interference by the Secretary of Health and Human Services with the negotiations between drug manufacturers and pharmacies and prescription drug plan sponsors. Grants the Secretary authority to negotiate contracts with manufacturers of covered Medicare part D (Voluntary Prescription Drug Benefit Program) drugs, consistent with requirements and in furtherance of the goals of providing quality care and containing costs under Medicare part D. Requires the Secretary to exercise such authority with respect to covered part D drugs offered under fallback prescription drug plans, if the Secretary determines that the negotiated prices are not fair and affordable prices compared to prices obtained by other Federal programs for such drugs. Eliminates the MA Regional Plan Stabilization Fund (Slush Fund). Directs the Secretary to ensure that payments to Medicare Advantage organizations are adjusted based on factors to ensure that the health status of the enrollee is reflected in such adjusted payments, including adjusting for the difference between the health status of the enrollee and individuals enrolled under the original Medicare fee-for-service program under parts A and B of title XVIII (Medicare) of the Social Security Act. Amends SSA title XVIII to provide for a waiver of the part D late enrollment penalty for the transition period
\\ \hline
ROUGEL-F1 & 
28.15
\\ \hline
\end{tabular}
\caption{Generated summaries by  (\gnews) on BillSum sampled by ROUGEL-F1.}
\end{table*}
 
















































\end{document}
