We performed experiments on the three datasets that correspond to a ``fair'' classification scenario and were previously used by~\cite{zemel2013learning}. In these datasets the ``nuisance'' or sensitive variable  is significantly correlated with the label  thus making the proper removal of  challenging. Furthermore, we also experimented with the Amazon reviews dataset to make a connection with the ``domain-adaptation'' literature. Finally, we also experimented with a more general task on the extended Yale B dataset; that of learning invariant representations. 

\subsection{Datasets}
For the fairness task we experimented with three datasets that were previously used by~\citet{zemel2013learning}. The German dataset is the smallest one with  data points and the objective is to predict whether a person has a good or bad credit rating. The sensitive variable is the gender of the individual. The Adult income dataset contains  entries and describes whether an account holder has over 50,000147,473p_\theta(\*x_n|{\*z_1}_n,\*s_n) = \text{Bern}(\*x_n | \!\pi_n = \sigma(f_\theta({\*z_1}_n, \*s_n)))\sigma(\cdot)\sigma(t) = \frac{1}{1 + e^{-t}}\*y\*xp_\theta(\*x_n|{\*z_1}_n, \*s_n) = \text{Poisson}(\*x_n | \!\lambda_n = e^{f_\theta({\*z_1}_n, \*s_n)})\*s\*yp(\*x_n|{\*z_1}_n, \*s_n)p_\theta(\*x_n|{\*z_1}_n, \*s_n) = \mathcal{N}(\*x_n| \!\mu_n = \sigma(f_\theta({\*z_1}_n, \*s_n)), \!\sigma_n = e^{f_\theta({\*z_1}_n, \*s_n)})\*z_1\*z_2\*z_1\*x\*z_1\*x\*z_2\*z_1\*z_1\*x\*z_2\*z_1\*z_1\*z_2q_\phi(\*y|\*z_1)\beta\alpha = 1 \*s\*y\alpha = 100 \cdot \frac{N_{\text{batch\_source}} + N_{\text{batch\_target}}}{N_{\text{batch\_source}}}\alpha = 200\beta = 100 \cdot N_{\text{batch}}\beta = 200 \cdot N_{\text{batch}}\*s\*y\*s\*s\*z_1\*s\*s\*s\*z_1\*yq_\phi(\*y|\*z_1)\*xq_\phi(\*z_1|\*x, \*s)K=50\*y\*s\*y\*s\*s\*y\*y\*s\*s\*s\*y\*y\*s\*z_1\*s\*x\*z_1\*z_1\*s\*x\*z_1\*s\*z_1\*s\*z_1\*s\*y\*y\*y\*x\*s\*x\*s\*y\*y\*s\*y\*s\*x\*x\*z_1$  from VFAE. Each example is plotted with the person ID and the image. Zoom in to see details.}
    \label{fig:yaleb_via}
\end{figure}
