@STRING{AG = "Aspect Graphs"}
@STRING{AV = "Active Vision"}

@STRING{CV = "Computer Vision"}
@STRING{Co = "Color"}

@STRING{Fa = "Face"}
@STRING{Fi = "Filter"}
@string{Ga = "Gabor"}
@string{Hi = "Histogram"}

@string{OR = "Object Recognition"}
@STRING{Refl = "Reflection and illumination"}

@string{Te = "Texture"}
@STRING{Tr = "Tracking"}


@STRING{TPAMI = "PAMI"}
@STRING{PAMI  = "PAMI"}
@STRING{IJCV  = "IJCV"}
@STRING{CVIU  = "CVIU"}
@STRING{CVGIP = "CVGIP:IU"}
@STRING{MVA   = "MVA"}
@STRING{PR    = "Pattern Recognition"}
@STRING{PRL   = "Pattern Recognition Letters"}
@STRING{IVC   = "Image and Vision Computing"}
@STRING{ML    = "Machine Learning"}

@STRING{FTCGV = "Foundations Trends Comp. Graphics Vision"}

@STRING{JSC    = "J. Sci. Comput."}

@STRING{JCN    = "J. Cog. Neurosci."}
@STRING(JEPHPP = "J. Exp. Psych.: Human Perception and Performance")
@STRING{JECP   = "J. Exp. Child Psych."}
@STRING{BBS    = "Behav. Brain Sc."}
@STRING{VR     = "Vis. Res."}
@STRING{AI     = "Art. Intell."}
@STRING{SIGMOD = "ACM SIGMOD"}

@STRING{CGA      = "Computer Graphics \& Applications"}
@STRING{PRESENCE = "Presence"}
@STRING{TCHI     = "ACM Trans. CHI"}
@STRING{CACM     = "Comm. ACM"}
@STRING{TACM     = "Trans. of the ACM"}
@STRING{TC       = "Trans. Comm."}
@STRING{TSE      = "IEEE Trans. Softw. Engin."}
@STRING{TMS      = "IEEE Trans. Math. Softw."}
@STRING{TIP      = "Trans. Image Proc."}
@STRING{TIS      = "IEICE Trans. Inf. & Syst."}
@STRING{TAC      = "IEEE Trans. Automatic Control"}
@STRING{TITS     = "TITS"}
@STRING{TVT      = "IEEE Trans. Vehic. Techn."}
@STRING{JRS      = "J. Robotic Systems"}
@STRING{AR       = "Autonomous Robots"}

@STRING{TOG      = "ACM Trans. Graphics"}

@STRING{ANNSTAT = "Annals of Statistics"}
@STRING{JC      = "J. Classif."}
@STRING{CAD     = "Cahiers de l'Analyse des Donn\'ees"}
@STRING{JMLR    = "JMLR"}

@STRING{JOE = "IEEE J. Oceanic Engineering"}

@STRING{JIVP = "JVIP"}

@STRING{SPL = "IEEE Signal Processing Letters"}

@STRING{IJCSI = "IJCSI"}


@STRING{ICCV   = "ICCV"}
@STRING{ICCV87 = "ICCV'87"}
@STRING{ICCV88 = "ICCV'88"}
@STRING{ICCV90 = "ICCV'90"}
@STRING{ICCV93 = "ICCV'93"}
@STRING{ICCV95 = "ICCV'95"}
@STRING{ICCV98 = "ICCV'98"}
@STRING{ICCV99 = "ICCV'99"}
@STRING{ICCV01 = "ICCV'01"}
@STRING{ICCV03 = "ICCV'03"}
@STRING{ICCV05 = "ICCV'05"}
@STRING{ICCV07 = "ICCV'07"}
@STRING{ICCV09 = "ICCV'09"}

@STRING{ECCV   = "ECCV"}
@STRING{ECCV90 = "ECCV'90"}
@STRING{ECCV92 = "ECCV'92"}
@STRING{ECCV94 = "ECCV'94"}
@STRING{ECCV96 = "ECCV'96"}
@STRING{ECCV98 = "ECCV'98"}
@STRING{ECCV00 = "ECCV'00"}
@STRING{ECCV02 = "ECCV'02"}
@STRING{ECCV04 = "ECCV'04"}
@STRING{ECCV06 = "ECCV'06"}
@STRING{ECCV08 = "ECCV'08"}
@STRING{ECCV10 = "ECCV'10"}

@string{CVPR   = "CVPR"}
@string{CVPR88 = "CVPR'88"}
@string{CVPR89 = "CVPR'89"}
@string{CVPR92 = "CVPR'92"}
@string{CVPR96 = "CVPR'96"}
@string{CVPR97 = "CVPR'97"}
@string{CVPR98 = "CVPR'98"}
@string{CVPR99 = "CVPR'99"}
@string{CVPR00 = "CVPR'00"}
@string{CVPR01 = "CVPR'01"}
@string{CVPR03 = "CVPR'03"}
@string{CVPR04 = "CVPR'04"}
@string{CVPR05 = "CVPR'05"}
@string{CVPR06 = "CVPR'06"}
@string{CVPR07 = "CVPR'07"}
@string{CVPR08 = "CVPR'08"}
@string{CVPR09 = "CVPR'09"}
@string{CVPR10 = "CVPR'10"}

@string{VPCVPR06 = "CVPR'06 Video Proceedings"}

@string{ACCV   = "ACCV"}
@string{ACCV06 = "ACCV'06"}
@string{ACCV07 = "ACCV'07"}

@string{ICPR   = "ICPR"}
@string{ICPR92 = "ICPR'92"}
@string{ICPR96 = "ICPR'96"}
@string{ICPR98 = "ICPR'98"}
@string{ICPR02 = "ICPR'02"}
@string{ICPR04 = "ICPR'04"}
@string{ICPR05 = "ICPR'05"}
@string{ICPR06 = "ICPR'06"}
@string{ICPR07 = "ICPR'07"}
@string{ICPR08 = "ICPR'08"}
@string{ICPR09 = "ICPR'09"}

@string{ICPRFGW = "ICPR'04 FG Net Workshop"}

@string{BMVC   = "BMVC"}
@string{BMVC91 = "BMVC'91"}
@string{BMVC98 = "BMVC'98"}
@string{BMVC99 = "BMVC'99"}
@string{BMVC00 = "BMVC'00"}
@string{BMVC01 = "BMVC'01"}
@string{BMVC02 = "BMVC'02"}
@string{BMVC03 = "BMVC'03"}
@string{BMVC04 = "BMVC'04"}
@string{BMVC05 = "BMVC'05"}
@string{BMVC06 = "BMVC'06"}
@string{BMVC07 = "BMVC'07"}
@string{BMVC08 = "BMVC'08"}
@string{BMVC09 = "BMVC'09"}
@string{BMVC10 = "BMVC'10"}

@string{BMCV02 = "Int'l Wksp. Biol. Motiv. Comp. Vis. (BMCV'02)"}

@string{DAGM   = "DAGM"}
@string{DAGM02 = "DAGM'02"}
@string{DAGM03 = "DAGM'03"}
@string{DAGM04 = "DAGM'04"}
@string{DAGM05 = "DAGM'05"}
@string{DAGM06 = "DAGM'06"}
@string{DAGM07 = "DAGM'07"}
@string{DAGM08 = "DAGM'08"}
@string{DAGM09 = "DAGM'09"}
@string{DAGM10 = "DAGM'10"}
@string{GCPR = "GCPR"}


@string{TDPVT = "3DPVT"}
@string{TDPVT06 = "3DPVT'06"}

@string{TDIM09 = "3DIM'09"}

@string{TDIMPVT = "3DIMPVT"}

@string{CAIP =   "International Conference on Analysis of Images
                                    and Patterns"}
@string{CAIP97 = "CAIP'97"}

@string{IJCAI77 = "IJCAI'77"}

@string{IJCAI = "IJCAI"}

@string{NIPS   = "NeurIPS"}
@string{NIPS97 = "NIPS'97"}
@string{NIPS01 = "NIPS'01"}
@string{NIPS02 = "NIPS'02"}
@string{NIPS03 = "NIPS'02"}
@string{NIPS06 = "NIPS'06"}

@string{ICML01 = "ICML'01"}
@string{ICML06 = "ICML'06"}

@string{ICIAP94 = "ICIAP'94"}
@string{ICIAP97 = "ICIAP'97"}

@string{ICIAR04 = "ICIAR'04"}

@string{CDC80 = "IEEE Conf. on Decision and Control"}

@string{SLCV04 = "ECCV'04 Workshop on Stat. Learn. in Comp. Vis."}

@string{TDRR = "ICCV Workshop on 3D Repr. and Recognition"}
@string{TDRR07 = "ICCV'07 Workshop on 3D Repr. for Recognition"}

@string{ICCVGIP08 = "ICCVGIP'08"}

@string{ICCVW09 = "ICCV Workshops"}

@string{POCV04 = "Workshop on Perc. Org. in Comp. Vis."}

@string{GMBV04 = "Workshop on Gener.-Model Based Vis."}

@string{WIVR04 = "Workshop on Image and Video Registr."}

@string{RMLE10 = "CVPR'09 Internet Vision Workshop"}

@string{RMLE10 = "ECCV'10 RMLE Workshop"}

@string{CVWW04 = "Comp. Vis. Winter Workshop (CVWW'04)"}

@string{FG95 = "Int. Workshop on Automatic Face and Gesture Recognition
(IWAFGR'95)"}
@string{FG96 = "Int. Conf. on Automatic Face and Gesture Recognition (FG'96)"}
@string{FG98 = "Int. Conf. on Automatic Face and Gesture Recognition (FG'98)"}
@string{FG = "Int. Conf. on Automatic Face and Gesture Recognition"}

@string{IROS   = "IROS"}
@string{IROS96 = "IROS'96"}

@string{ICRA   = "ICRA"}
@string{ICRA95 = "ICRA'95"}
@string{ICRA00 = "ICRA'00"}
@string{ICRA01 = "ICRA'01"}
@string{ICRA02 = "ICRA'02"}
@string{ICRA03 = "ICRA'03"}
@string{ICRA04 = "ICRA'04"}
@string{ICRA05 = "ICRA'05"}
@string{ICRA06 = "ICRA'06"}
@string{ICRA07 = "ICRA'07"}
@string{ICRA08 = "ICRA'08"}
@string{ICRA09 = "ICRA'09"}
@string{ICRA11 = "ICRA'11"}

@string{SIGGRAPH   = "SIGGRAPH"}
@string{SIGGRAPH94 = "SIGGRAPH'94"}
@string{SIGGRAPH97 = "SIGGRAPH'97"}
@string{SIGGRAPH99 = "SIGGRAPH'99"}
@string{SIGGRAPH03 = "SIGGRAPH'03"}
@string{SIGGRAPH04 = "SIGGRAPH'04"}
@string{SIGGRAPH05 = "SIGGRAPH'05"}
@string{SIGGRAPH06 = "SIGGRAPH'06"}
@string{SIGGRAPH07 = "SIGGRAPH'07"}
@string{SIGGRAPH08 = "SIGGRAPH'08"}
@string{SIGGRAPH09 = "SIGGRAPH'09"}
@string{SIGGRAPH10 = "SIGGRAPH'10"}

@string{CHI92 = "CHI'92"}
@string{CHI97 = "CHI'97"}

@string{ICIP   = "ICIP"}
@string{ICIP06 = "ICIP'06"}

@STRING{VR99  = "Virtual Reality '99"}
@STRING{VR00  = "Virtual Reality 2000"}

@STRING{UIST92 = "UIST'92"}
@STRING{UIST97 = "UIST'97"}

@STRING{WSCG99 = "WSCG'99"}

@STRING{ICASSP = "ICASSP"}

@string{SCIA93 = "SCIA'93"}
@string{SCIA91 = "SCIA'91"}

@string{EUROCOLT = "Europ. Conf. on Comput. Learning Theory"}

@string{IVS96 = {Intel. Vehicles Symp.'96}}
@string{IVS00 = {Intel. Vehicles Symp.'00}}
@string{IVS02 = {Intel. Vehicles Symp.'02}}
@string{IVS03 = {Intel. Vehicles Symp.'03}}
@string{IVS04 = {Intel. Vehicles Symp.'04}}
@string{IVS05 = {Intel. Vehicles Symp.'05}}
@string{IVS08 = {Intel. Vehicles Symp.'08}}
@string{IVS11 = {Intel. Vehicles Symp.'11}}

@string{WIT09 = {Int. Workshop Intel. Transp. (WIT'09)}}

@string{INES04 = {Intel. Eng. Syst. (INES'04)}}

@string{ICANN98 = {ICANN'98}}


@string{SEQS97 = {SEQS'97}}

@string{CIVR08 = "CIVR'08"}
@string{CIVR07 = "CIVR'07"}

@string{ICML = "ICML"}

@string{SPIE = "SPIE"}

@string{JFR = "Journal of Field Robotics"}

@string{PCVIA = "Photogrammetric Computer Vision and Image Analysis"}

@string{RSS = "RSS"}

@string{UAI = "UAI"}

@string{IVS = "Intel. Vehicles Symp."}

@string{ISRR = "ISRR"}

@string{ITSC = "ITSC"}

@string{AAAI = "AAAI"}

@string{RAS = "RAS"}

@string{IJRR = "IJRR"}

@string{WACV = "WACV"}

@string{ICME = "ICME"}

@string{ICLR = "ICLR"}

@string{IAS = "IAS"}

@string{ROMAN = "RO-MAN"}

@string{VCIP = "VCIP"}

@string{TCSVT = "TCSVT"}

@string{ISCAS = "ISCAS"}

@string{FutureTech = "FutureTech"}

@string{IJSR = "IJSR"}

@string{ACRA = "ACRA"}

@string{MICCAI = "MICCAI"}

@string{CSaE = "CS\&E"}

@string{RAL = "RA-L"}

@string{RAM = "RAM"}

@string{Sensors = "Sensors"}

@string{NIPSADW = "NIPS Autodiff Workshop"}

@string{RSSWDL = "RSS, Workshop on Limits and Potentials of Deep Learning in Robotics"}

@string{CoRL = "CoRL"}

@string{WACV = "WACV"}

@string{ICMR = "ICMR"}

@string{NeurIPS = "NeurIPS"}

@string{ACL = "ACL"}

@string{ThreeDV = "3DV"}