Zhiwu Qing:
a. Made a significant intellectual contribution to the implementation of whole experimental system including but not limited to the efficient implementation of ideas, the model construction, the reproduction of the codes and the deployment of the experiments.
b. Contributed to the preparation materials for the methodology part and writing the draft of the full manuscript.



Shiwei Zhang:
a. Made a significant intellectual contribution to the core ideas of the manuscript. Participating in the design of the analytical experiments, optimizing code implementation, and reproducing the results of the related works.
b. Contributed to the preparation materials for the related works, revising the introduction of the article.



Ziyuan Huang:
a. Made a significant intellectual contribution to the analysis of the experiments. Applying the statistical and mathematical techniques to analyze the experimental results, and investigating relevant theoretical explanations.
b. Contributed to the preparation materials for the experiments part, editing the figures and tables in the experiment part of the manuscript.


Xiang Wang:
a. Made a significant intellectual contribution to the investigation of the existing published works and software developments. Participating in the design of experiments, collecting the related pre-trained models and datasets.
b. Contributed to revising the experiments of the manuscript, reviewing the experiments of the manuscript.


Yuehuan Wang:
a. Made a significant intellectual contribution to the visualizations of the experimental results. Participating in the design of the experiments, making an in-depth theoretical analysis of the experimental results, and looking for the theoretical basis.
b. Contributed to revising the methodology of the manuscript.


Yiliang Lv:
a. Made a significant intellectual contribution to the guidance of the detailed experimental designs and experimental optimization. Addressing the demand for computational resources for large model experiments.
b. Contributed to revising the related works of the manuscript. 



Changxin Gao:
a. Made a significant intellectual contribution to the formulation of research goals and aims. Participating in the design of the system, reviewing and editing the figures in the methodology part of the manuscript and analyzing the experimental results;
b. Contributed to the critical review and commentary of the manuscript.


Nong Sang:
a. Made a significant intellectual contribution to the guidance of the theoretical basis of experimental results, and leadership and supervision responsibility for the research activity planning and
execution;
b. Contributed to the preparation materials for the introduction part, critical review and commentary of the full manuscript.

