\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfiles}
\usepackage{multirow}
\usepackage{booktabs}



\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\cvprfinalcopy 

\def\cvprPaperID{4322} \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\title{Vision-Language Navigation with Self-Supervised Auxiliary Reasoning Tasks}

\author{Fengda Zhu\hspace{6mm}
Yi Zhu\hspace{6mm}
Xiaojun Chang \hspace{6mm}
Xiaodan Liang \\
Monash University \hspace{2mm} University of Chinese Academy of Sciences \\
Sun Yat-sen University \hspace{2mm} Dark Matter AI Inc.  \\
{\tt\small zhufengda@yahoo.com \hspace{2mm} zhu.yee@outlook.com } \\
{\tt\small cxj273@gmail.com \hspace{2mm} xdliang328@gmail.com }
}


\maketitle
\thispagestyle{empty}



\subfile{abstract}
\subfile{introduction}
\subfile{related_work}
\subfile{method}
\subfile{experiment}

\clearpage

\begin{thebibliography}{10}\itemsep=-1pt
	
	\bibitem{agrawal2015vqa}
	Aishwarya {Agrawal}, Jiasen {Lu}, Stanislaw {Antol}, Margaret {Mitchell},
	C.~Lawrence {Zitnick}, Dhruv {Batra}, and Devi {Parikh}.
	\newblock Vqa: Visual question answering.
	\newblock {\em arXiv preprint arXiv:1505.00468}, 2015.
	
	\bibitem{alberti2019fusion}
	Chris {Alberti}, Jeffrey {Ling}, Michael {Collins}, and David {Reitter}.
	\newblock Fusion of detected objects in text for visual question answering.
	\newblock In {\em 2019 Conference on Empirical Methods in Natural Language
		Processing}, 2019.
	
	\bibitem{anderson2018on}
	Peter {Anderson}, Angel~X. {Chang}, Devendra~Singh {Chaplot}, Alexey
	{Dosovitskiy}, Saurabh {Gupta}, Vladlen {Koltun}, Jana {Kosecka}, Jitendra
	{Malik}, Roozbeh {Mottaghi}, Manolis {Savva}, and Amir~Roshan {Zamir}.
	\newblock On evaluation of embodied navigation agents.
	\newblock {\em arXiv preprint arXiv:1807.06757}, 2018.
	
	\bibitem{anderson2018bottom}
	Peter {Anderson}, Xiaodong {He}, Chris {Buehler}, Damien {Teney}, Mark
	{Johnson}, Stephen {Gould}, and Lei {Zhang}.
	\newblock Bottom-up and top-down attention for image captioning and visual
	question answering.
	\newblock In {\em 2018 IEEE/CVF Conference on Computer Vision and Pattern
		Recognition}, pages 6077--6086, 2018.
	
	\bibitem{anderson2018vision}
	Peter {Anderson}, Qi {Wu}, Damien {Teney}, Jake {Bruce}, Mark {Johnson}, Niko
	{Sunderhauf}, Ian {Reid}, Stephen {Gould}, and Anton van~den {Hengel}.
	\newblock Vision-and-language navigation: Interpreting visually-grounded
	navigation instructions in real environments.
	\newblock In {\em 2018 IEEE/CVF Conference on Computer Vision and Pattern
		Recognition}, pages 3674--3683, 2018.
	
	\bibitem{antol2015vqa}
	Stanislaw {Antol}, Aishwarya {Agrawal}, Jiasen {Lu}, Margaret {Mitchell}, Dhruv
	{Batra}, C.~Lawrence {Zitnick}, and Devi {Parikh}.
	\newblock Vqa: Visual question answering.
	\newblock In {\em 2015 IEEE International Conference on Computer Vision
		(ICCV)}, pages 2425--2433, 2015.
	
	\bibitem{bojarski2016end}
	Mariusz {Bojarski}, Davide~Del {Testa}, Daniel {Dworakowski}, Bernhard
	{Firner}, Beat {Flepp}, Prasoon {Goyal}, Lawrence~D. {Jackel}, Mathew
	{Monfort}, Urs {Muller}, Jiakai {Zhang}, Xin {Zhang}, Jake {Zhao}, and Karol
	{Zieba}.
	\newblock End to end learning for self-driving cars.
	\newblock {\em arXiv preprint arXiv:1604.07316}, 2016.
	
	\bibitem{chang2017matterport3d}
	Angel {Chang}, Angela {Dai}, Thomas {Funkhouser}, Maciej {Halber}, Matthias
	{Niebner}, Manolis {Savva}, Shuran {Song}, Andy {Zeng}, and Yinda {Zhang}.
	\newblock Matterport3d: Learning from rgb-d data in indoor environments.
	\newblock In {\em 2017 International Conference on 3D Vision (3DV)}, pages
	667--676, 2017.
	
	\bibitem{devlin2018bert}
	Jacob {Devlin}, Ming-Wei {Chang}, Kenton {Lee}, and Kristina {Toutanova}.
	\newblock Bert: Pre-training of deep bidirectional transformers for language
	understanding.
	\newblock {\em arXiv preprint arXiv:1810.04805}, 2018.
	
	\bibitem{fried2018speaker}
	Daniel {Fried}, Ronghang {Hu}, Volkan {Cirik}, Anna {Rohrbach}, Jacob
	{Andreas}, Louis-Philippe {Morency}, Taylor {Berg-Kirkpatrick}, Kate
	{Saenko}, Dan {Klein}, and Trevor {Darrell}.
	\newblock Speaker-follower models for vision-and-language navigation.
	\newblock In {\em NIPS 2018: The 32nd Annual Conference on Neural Information
		Processing Systems}, pages 3314--3325, 2018.
	
	\bibitem{gidaris2018unsupervised}
	Spyros {Gidaris}, Praveer {Singh}, and Nikos {Komodakis}.
	\newblock Unsupervised representation learning by predicting image rotations.
	\newblock In {\em ICLR 2018 : International Conference on Learning
		Representations 2018}, 2018.
	
	\bibitem{gu2019scene}
	Jiuxiang {Gu}, Handong {Zhao}, Zhe {Lin}, Sheng {Li}, Jianfei {Cai}, and
	Mingyang {Ling}.
	\newblock Scene graph generation with external knowledge and image
	reconstruction.
	\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
		Pattern Recognition}, pages 1969--1978, 2019.
	
	\bibitem{gupta2017cognitive}
	Saurabh {Gupta}, Varun {Tolani}, James {Davidson}, Sergey {Levine}, Rahul
	{Sukthankar}, and Jitendra {Malik}.
	\newblock Cognitive mapping and planning for visual navigation.
	\newblock {\em arXiv preprint arXiv:1702.03920}, 2017.
	
	\bibitem{he2015deep}
	Kaiming {He}, Xiangyu {Zhang}, Shaoqing {Ren}, and Jian {Sun}.
	\newblock Deep residual learning for image recognition.
	\newblock {\em arXiv preprint arXiv:1512.03385}, 2015.
	
	\bibitem{ho2016generative}
	Jonathan {Ho} and Stefano {Ermon}.
	\newblock Generative adversarial imitation learning.
	\newblock {\em arXiv preprint arXiv:1606.03476}, 2016.
	
	\bibitem{jaderberg2017reinforcement}
	Max {Jaderberg}, Volodymyr {Mnih}, Wojciech~Marian {Czarnecki}, Tom {Schaul},
	Joel~Z {Leibo}, David {Silver}, and Koray {Kavukcuoglu}.
	\newblock Reinforcement learning with unsupervised auxiliary tasks.
	\newblock In {\em ICLR 2017 : International Conference on Learning
		Representations 2017}, 2017.
	
	\bibitem{kempka2016vizdoom}
	Michał {Kempka}, Marek {Wydmuch}, Grzegorz {Runc}, Jakub {Toczek}, and
	Wojciech {Jaśkowski}.
	\newblock Vizdoom: A doom-based ai research platform for visual reinforcement
	learning.
	\newblock {\em arXiv preprint arXiv:1605.02097}, 2016.
	
	\bibitem{kolve2017ai2}
	Eric {Kolve}, Roozbeh {Mottaghi}, Daniel {Gordon}, Yuke {Zhu}, Abhinav {Gupta},
	and Ali {Farhadi}.
	\newblock Ai2-thor: An interactive 3d environment for visual ai.
	\newblock {\em arXiv preprint arXiv:1712.05474}, 2017.
	
	\bibitem{lan2019albert}
	Zhenzhong {Lan}, Mingda {Chen}, Sebastian {Goodman}, Kevin {Gimpel}, Piyush
	{Sharma}, and Radu {Soricut}.
	\newblock Albert: A lite bert for self-supervised learning of language
	representations.
	\newblock {\em arXiv preprint arXiv:1909.11942}, 2019.
	
	\bibitem{liu2019self}
	Shikun {Liu}, Andrew {Davison}, and Edward {Johns}.
	\newblock Self-supervised generalisation with meta auxiliary learning.
	\newblock In {\em NeurIPS 2019 : Thirty-third Conference on Neural Information
		Processing Systems}, 2019.
	
	\bibitem{lu2019vilbert}
	Jiasen {Lu}, Dhruv {Batra}, Devi {Parikh}, and Stefan {Lee}.
	\newblock Vilbert: Pretraining task-agnostic visiolinguistic representations
	for vision-and-language tasks.
	\newblock In {\em NeurIPS 2019 : Thirty-third Conference on Neural Information
		Processing Systems}, 2019.
	
	\bibitem{lu2016hierarchical}
	Jiasen {Lu}, Jianwei {Yang}, Dhruv {Batra}, and Devi {Parikh}.
	\newblock Hierarchical question-image co-attention for visual question
	answering.
	\newblock In {\em Advances in Neural Information Processing Systems}, pages
	289--297, 2016.
	
	\bibitem{ma2019self}
	Chih-Yao {Ma}, jiasen {lu}, Zuxuan {Wu}, Ghassan {AlRegib}, Zsolt {Kira},
	richard {socher}, and Caiming {Xiong}.
	\newblock Self-monitoring navigation agent via auxiliary progress estimation.
	\newblock In {\em ICLR 2019 : 7th International Conference on Learning
		Representations}, 2019.
	
	\bibitem{mikolov2013distributed}
	Tomas {Mikolov}, Ilya {Sutskever}, Kai {Chen}, Greg~S {Corrado}, and Jeff
	{Dean}.
	\newblock Distributed representations of words and phrases and their
	compositionality.
	\newblock In {\em Advances in Neural Information Processing Systems 26}, pages
	3111--3119, 2013.
	
	\bibitem{mirowski2017learning}
	Piotr {Mirowski}, Razvan {Pascanu}, Fabio {Viola}, Hubert {Soyer}, Andy
	{Ballard}, Andrea {Banino}, Misha {Denil}, Ross {Goroshin}, Laurent {Sifre},
	Koray {Kavukcuoglu}, Dharshan {Kumaran}, and Raia {Hadsell}.
	\newblock Learning to navigate in complex environments.
	\newblock In {\em ICLR 2017 : International Conference on Learning
		Representations 2017}, 2017.
	
	\bibitem{mnih2016asynchronous}
	Volodymyr {Mnih}, Adrià~Puigdomènech {Badia}, Mehdi {Mirza}, Alex {Graves},
	Tim {Harley}, Timothy~P. {Lillicrap}, David {Silver}, and Koray
	{Kavukcuoglu}.
	\newblock Asynchronous methods for deep reinforcement learning.
	\newblock In {\em ICML'16 Proceedings of the 33rd International Conference on
		International Conference on Machine Learning - Volume 48}, pages 1928--1937,
	2016.
	
	\bibitem{odena2017conditional}
	Augustus {Odena}, Christopher {Olah}, and Jonathon {Shlens}.
	\newblock Conditional image synthesis with auxiliary classifier gans.
	\newblock In {\em ICML'17 Proceedings of the 34th International Conference on
		Machine Learning - Volume 70}, pages 2642--2651, 2017.
	
	\bibitem{pathak2017curiosity}
	Deepak {Pathak}, Pulkit {Agrawal}, Alexei~A. {Efros}, and Trevor {Darrell}.
	\newblock Curiosity-driven exploration by self-supervised prediction.
	\newblock In {\em International Conference on Machine Learning}, pages
	2778--2787, 2017.
	
	\bibitem{pearl1988probabilistic}
	Judea {Pearl}.
	\newblock {\em Probabilistic Reasoning in Intelligent Systems: Networks of
		Plausible Inference}.
	\newblock 1988.
	
	\bibitem{racanire2017imagination}
	Sébastien {Racanière}, Theophane {Weber}, David~P. {Reichert}, Lars
	{Buesing}, Arthur {Guez}, Danilo~Jimenez {Rezende}, Adrià~Puigdomènech
	{Badia}, Oriol {Vinyals}, Nicolas {Heess}, Yujia {Li}, Razvan {Pascanu},
	Peter~W. {Battaglia}, Demis {Hassabis}, David {Silver}, and Daan {Wierstra}.
	\newblock Imagination-augmented agents for deep reinforcement learning.
	\newblock In {\em Advances in Neural Information Processing Systems}, pages
	5690--5701, 2017.
	
	\bibitem{ren2015faster}
	Shaoqing {Ren}, Kaiming {He}, Ross {Girshick}, and Jian {Sun}.
	\newblock Faster r-cnn: Towards real-time object detection with region proposal
	networks.
	\newblock {\em arXiv preprint arXiv:1506.01497}, 2015.
	
	\bibitem{schulman2017proximal}
	John {Schulman}, Filip {Wolski}, Prafulla {Dhariwal}, Alec {Radford}, and Oleg
	{Klimov}.
	\newblock Proximal policy optimization algorithms.
	\newblock {\em arXiv preprint arXiv:1707.06347}, 2017.
	
	\bibitem{su2019vl}
	Weijie {Su}, Xizhou {Zhu}, Yue {Cao}, Bin {Li}, Lewei {Lu}, Furu {Wei}, and
	Jifeng {Dai}.
	\newblock Vl-bert: Pre-training of generic visual-linguistic representations.
	\newblock {\em arXiv preprint arXiv:1908.08530}, 2019.
	
	\bibitem{sun2019unsupervised}
	Yu {Sun}, Eric {Tzeng}, Trevor {Darrell}, and Alexei~A. {Efros}.
	\newblock Unsupervised domain adaptation through self-supervision.
	\newblock {\em arXiv preprint arXiv:1909.11825}, 2019.
	
	\bibitem{sun2019test}
	Yu {Sun}, Xiaolong {Wang}, Zhuang {Liu}, John {Miller}, Alexei~A. {Efros}, and
	Moritz {Hardt}.
	\newblock Test-time training for out-of-distribution generalization.
	\newblock {\em arXiv preprint arXiv:1909.13231}, 2019.
	
	\bibitem{tan2019lxmert}
	Hao {Tan} and Mohit {Bansal}.
	\newblock Lxmert: Learning cross-modality encoder representations from
	transformers.
	\newblock In {\em 2019 Conference on Empirical Methods in Natural Language
		Processing}, 2019.
	
	\bibitem{tan2019learning}
	Hao {Tan}, Licheng {Yu}, and Mohit {Bansal}.
	\newblock Learning to navigate unseen environments: Back translation with
	environmental dropout.
	\newblock In {\em NAACL-HLT 2019: Annual Conference of the North American
		Chapter of the Association for Computational Linguistics}, pages 2610--2621,
	2019.
	
	\bibitem{thomason2019vision}
	Jesse {Thomason}, Michael {Murray}, Maya {Cakmak}, and Luke {Zettlemoyer}.
	\newblock Vision-and-dialog navigation.
	\newblock {\em arXiv preprint arXiv:1907.04957}, 2019.
	
	\bibitem{veeriah2019discovery}
	Vivek {Veeriah}, Matteo {Hessel}, Zhongwen {Xu}, Richard {Lewis}, Janarthanan
	{Rajendran}, Junhyuk {Oh}, Hado van {Hasselt}, David {Silver}, and Satinder
	{Singh}.
	\newblock Discovery of useful questions as auxiliary tasks.
	\newblock {\em arXiv preprint arXiv:1909.04607}, 2019.
	
	\bibitem{veeriah2019the}
	Vivek {Veeriah}, Richard~L {Lewis}, Janarthanan {Rajendran}, David {Silver},
	and Satinder {Singh}.
	\newblock The discovery of useful questions as auxiliary tasks.
	\newblock In {\em NeurIPS 2019 : Thirty-third Conference on Neural Information
		Processing Systems}, 2019.
	
	\bibitem{wang2018reinforced}
	Xin {Wang}, Qiuyuan {Huang}, Asli {Celikyilmaz}, Jianfeng {Gao}, Dinghan
	{Shen}, Yuan-Fang {Wang}, William~Yang {Wang}, and Lei {Zhang}.
	\newblock Reinforced cross-modal matching and self-supervised imitation
	learning for vision-language navigation.
	\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
		Pattern Recognition}, pages 6629--6638, 2018.
	
	\bibitem{wang2018look}
	Xin {Wang}, Wenhan {Xiong}, Hongmin {Wang}, and William~Yang {Wang}.
	\newblock Look before you leap: Bridging model-free and model-based
	reinforcement learning for planned-ahead vision-and-language navigation.
	\newblock {\em arXiv preprint arXiv:1803.07729}, 2018.
	
	\bibitem{wu2018building}
	Yi {Wu}, Yuxin {Wu}, Georgia {Gkioxari}, and Yuandong {Tian}.
	\newblock Building generalizable agents with a realistic and rich 3d
	environment.
	\newblock In {\em ICLR 2018 : International Conference on Learning
		Representations 2018}, 2018.
	
	\bibitem{zellers2018from}
	Rowan {Zellers}, Yonatan {Bisk}, Ali {Farhadi}, and Yejin {Choi}.
	\newblock From recognition to cognition: Visual commonsense reasoning.
	\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
		Pattern Recognition}, pages 6720--6731, 2018.
	
	\bibitem{zhao2017pyramid}
	Hengshuang {Zhao}, Jianping {Shi}, Xiaojuan {Qi}, Xiaogang {Wang}, and Jiaya
	{Jia}.
	\newblock Pyramid scene parsing network.
	\newblock In {\em 2017 IEEE Conference on Computer Vision and Pattern
		Recognition (CVPR)}, pages 6230--6239, 2017.
		
	\bibitem{zeng2020dense}
	Runhao {Zeng}, Haoming {Xu}, Wenbing {Huang}, Peihao {Chen}, Mingkui
	{Tan}, and Chuang {Gan}.
	\newblock Dense Regression Network for Video Grounding.
	\newblock In {\em 2020 IEEE Conference on Computer Vision and Pattern
		Recognition (CVPR)}, 2020.
	
	\bibitem{zhu2020vision}
	Yi {Zhu}, Fengda {Zhu}, Zhaohuan {Zhan}, Bingqian {Lin}, Jianbin
	{Jiao},  Xiaojun {Chang}, and Xiaodan {Liang}.
	\newblock Vision-Dialog Navigation by Exploring Cross-modal Memory.
	\newblock In {\em 2020 IEEE Conference on Computer Vision and Pattern
		Recognition (CVPR)}, 2020.
	
	\bibitem{ma2019regretful}
	Chih-Yao {Ma}, Zuxuan {Wu}, Ghassan {AlRegib}, Caiming {Xiong}, and Zsolt
	{Kira}.
	\newblock The regretful agent: Heuristic-aided navigation through progress estimation.
	\newblock In {\em 2019 IEEE Conference on Computer Vision and Pattern
		Recognition (CVPR)}, pages 6732--6740, 2019.
	
	\bibitem{ke2019tactical}
	Liyiming {Ke}, Xiujun {Li}, Yonatan {Bisk}, Ari {Holtzman}, Zhe
	{Gan}, Jingjing {Liu}, Jianfeng {Gao}, Yejin {Choi}, and Siddhartha {Srinivasa}.
	\newblock Tactical rewind: Self-correction via backtracking in vision-and-language navigation.
	\newblock In {\em 2019 IEEE Conference on Computer Vision and Pattern
		Recognition (CVPR)}, pages 6741--6749, 2019.
	
	\bibitem{li2019robust}
	Xiujun {Li}, Chunyuan {Li}, Qiaolin {Xia}, Yonatan {Bisk}, Asli
	{Celikyilmaz}, Jianfeng {Gao}, Noah {Smith}, and Yejin {Choi}.
	\newblock Robust Navigation with Language Pretraining and Stochastic Sampling.
	\newblock {\em arXiv preprint arXiv:1909.02244}, 2019.
	
	\bibitem{huang2019transferable}
	Haoshuo {Huang}, Vihan {Jain}, Harsh {Mehta}, Alexander {Ku}, Gabriel
	{Magalhaes}, Jason {Baldridge}, and Eugene {Ie}.
	\newblock Transferable representation learning in vision-and-language navigation.
	\newblock In {\em 2019 IEEE International Conference on Computer Vision (ICCV)}, pages 7404--7413, 2019.
	
\end{thebibliography}


\end{document}
